---
title: "C-center models"
output: 
  html_document: 
    number_sections: yes
    toc: yes
    toc_float: yes
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load required packages
```{r}
library(tidyr)
library(dplyr)
library(stringr)
library(lme4)
library(car)
library(visreg)
library(optimx)
library(lmerTest)
library(emmeans)
library(neldermead)
library(ggplot2)
library(optimx)
library(mgcv)
library(itsadug)
```

# Introduction
From the c-center hypothesis it follows that onsets are coordinated inphase with the following vowel, meaning that they are initiated roughly around the same time. On the other hand, multiple consonants in an onset are also coordinated antiphase with each other (i.e. sequential initiation). To cope with these competing relations, the leftmost C gets shifted towards the left and the rightmost C gets shifted towards the right (with respect to the following vowel) in complex onsets. In this way, the average onset of the two C's is still initiated at the same time as the vowel, whereas they are also still coordinated in an antiphase relation with each other. Our hypothesis is that PD patients will show a stronger preference for the inphase relationship, coming at the cost of the antiphase relationship. In our data we've included three singular onsets /p/, /m/ and /x/ and four complex onsets: /sp/, /sm/, /sx/ and /spr/. The variable 'time.lag.norm' is the normalized time lag between the offset of the leftmost C and the onset of the rightmost C. If our hypothesis is correct the lag between the C's should be lower for the PD group in comparison with the controls.


## Read data
```{r save data}
df <- readRDS("../data/modelling_data_CC.rds")
```

## Clean data
Remove all the rows that have NA in the dependent variable and drop 'empty' levels, convert variables to factors and create trimmed dataset.
```{r remove NA}
df$subj <- as.factor(df$subj)

df <- df[complete.cases(df$time.lag.norm),]
df <- droplevels(df)

# Sort data
df <- df[order(df$subj, df$recording.no),]

# convert to factors/numeric
df$prompt <- as.factor(df$prompt)
df$condition <- as.factor(df$condition)

# remove all the recordings without a speech rate measure:
# df <- df %>% drop_na(duration.tt)

# remove participants with only a few datapoints
#remove <- c("PD19", "CTRL19")
#df.trim <- df[! df$subj %in% remove, ]
#df.trim <- droplevels(df.trim)


# create df without sx
#df.trim <- df.trim[! df.trim$cluster == "sx", ]
#df.trim <- droplevels(df.trim)
#levels(df.trim$cluster)

# create df without spr
df.nospr <- df[! df$cluster == "spr", ]
df.nospr <- droplevels(df.nospr)
levels(df.nospr$cluster)
df.nospr$cluster <- relevel(df.nospr$cluster, ref = "sm")



# merge /sm/ and /sp/
df.nospr$poa <- ifelse(df.nospr$cluster == "sx","C2lingual","C2bilabial")
df.nospr$poa <- as.factor(df.nospr$poa)
table(df.nospr$poa, df.nospr$cluster)


```


# LME models for CC lag analysis (coarticulation)
If indeed patients would show a stronger preference for inphase coordination, this would mean that they would produce the C gestures in the syllable's onset closer together in time. For this, we look at the lag between the gestural offset of the leftmost C and the gestural onset of the righmost C in complex clusters.

## Hypothesis testing

Fitting first model, smooth over subject:
```{r}
summary(m <- bam(time.lag.norm ~ s(subj, bs='re'), data=df.nospr, method = "ML"))
```

Adding group as fixed effect:
```{r}
summary(m1 <- bam(time.lag.norm ~ s(subj,bs='re') + group, data=df.nospr, method = "ML"))
compareML(m,m1)
```
Sign. improvement.

### Exploratory analysis


Adding cluster as fixed effect:
```{r}
summary(m2 <- bam(time.lag.norm ~ s(subj,bs='re') + cluster, data=df.nospr, method = "ML"))
compareML(m1,m2)
```
Improves.

Adding cluster:group as fixed effect:
```{r}
summary(m3 <- bam(time.lag.norm ~ s(subj,bs='re') + cluster*group, data=df.nospr, method = "ML"))
compareML(m2,m3)
```
Sign. improvement.


Add by-subj random slopes for cluster
```{r}
summary(m3 <- bam(time.lag.norm ~ s(subj,bs='re') + cluster*group, data=df.nospr, method="REML")) #fit m3 using REML

summary(m4 <- bam(time.lag.norm ~ s(subj,bs='re') + s(subj, cluster, bs="re") + cluster*group, data=df.nospr, method="REML"))
compareML(m3, m4)
```
We need the random slope.



Add speech rate.
```{r}
m4 <- bam(time.lag.norm ~ s(subj,bs='re') + s(subj, cluster, bs="re") + cluster*group, data=df.nospr, method = "ML") #fit using ML

df.nospr$duration.tt <- df.nospr$duration.tt - mean(df.nospr$duration.tt) # center continues variable

summary(m5 <- bam(time.lag.norm ~ s(subj,bs='re') + s(subj, cluster, bs="re") + cluster*group + s(duration.tt), data=df.nospr, method = "ML"))
compareML(m4, m5)
```
Not sign at alpha = 0.010.

Gender
```{r}
summary(m6 <- bam(time.lag.norm ~ s(subj,bs='re') + s(subj, cluster, bs="re") + cluster*group + gender, data=df.nospr, method = "ML"))
compareML(m4, m6)
```
No sign. improvement.

Add recording.no
```{r}
summary(m7 <- bam(time.lag.norm ~ s(subj,bs='re') + s(subj, cluster, bs="re") + cluster*group + s(recording.no), data=df.nospr, method = "ML"))
compareML(m4, m7, suggest.report=T)
```
No improvement.

Add age.
```{r}
summary(m8 <- bam(time.lag.norm ~ s(subj,bs='re') + s(subj, cluster, bs="re") + cluster*group + s(age), data=df.nospr, method = "ML"))
compareML(m4, m8)
```
No improvement.

Add dialect.
```{r}
summary(m9 <- bam(time.lag.norm ~ s(subj,bs='re') + s(subj, cluster, bs="re") + cluster*group + dialect, data=df.nospr, method = "ML"))
compareML(m4, m9)
```
No improvement.


#### Visualize final model:
```{r}
summary(m4REML <- bam(time.lag.norm ~ s(subj,bs='re') + s(subj, cluster, bs="re") + cluster*group, data=df.nospr, method = 'REML')) # refit using REML


summary(m4_scat <- bam(time.lag.norm ~ s(subj,bs='re') + s(subj, cluster, bs="re") + cluster*group, data=df.nospr, method = 'fREML', family = "scat", discrete = T)) # refit using REML

#residuals niet normaal verdeeld, ook gefit met scat, zelfde patroon

summary(m6.lme <- lmer(time.lag.norm ~ cluster*group+ (1+cluster|subj), data=df.nospr, REML=F)) # refit using REML

visreg(m4REML, "cluster", by = "group", overlay = T, points = list(cex = 0.1), ylab = "Inter plateau interval (ms)", xlab = "Cluster")

gam.check(m4REML)
library("car")
qqp(resid(m4REML))
plot(fitted(m4REML),resid(m4REML))

```


### Check effect of tremor and akinetic on time.lag.norm for PD group only
Base model:
```{r}
summary(m6.PD <- bam(time.lag.norm ~ s(subj,bs='re') + s(subj, cluster, bs="re") + cluster + s(duration.tt), data=df.nospr[df.nospr$group == "PD",], method = "ML"))
```

```{r}
summary(m7.PD <- bam(time.lag.norm ~ s(subj,bs='re') + s(subj, cluster, bs="re") + cluster + s(duration.tt) + tremor, data=df.nospr[df.nospr$group == "PD",], method = "ML"))
compareML(m6.PD, m7.PD)
```
No improvement.

Add akinetic as a fixed effect.
```{r}
summary(m8.PD <- bam(time.lag.norm ~ s(subj,bs='re') + s(subj, cluster, bs="re") + cluster + s(duration.tt) + akinetic, data=df.nospr[df.nospr$group == "PD",], method = "ML"))
compareML(m6.PD, m8.PD)
```
No improvement.

# Check correlations C1/C2 with IPI

## Hypothesis testing
Create one var for duration of C1/C2
```{r}
df.nospr <- df.nospr %>%
  gather(whichC, durationC, dur.C1.norm, dur.C2.norm) %>%
  arrange(subj, group, recording.no)

df.nospr$whichC <- as.factor(df.nospr$whichC)

df.nospr$durationC <- df.nospr$durationC - mean(df.nospr$durationC) # center continues variable
```



Base model.
```{r}
summary(durC_m <- bam(durationC ~ s(subj,bs='re') + s(subj, whichC, bs="re"), data = df.nospr, method = "ML"))
```

```{r}
summary(durC_m1 <- bam(durationC ~ group + s(subj,bs='re') + s(subj, whichC, bs="re"), data = df.nospr, method = "ML"))
compareML(durC_m, durC_m1)
```
No diff.

### Exploratory analysis


Add C duration
```{r}
summary(durC_m2 <- bam(durationC ~ whichC + s(subj,bs='re') + s(subj, whichC, bs="re"), data = df.nospr, method = "ML"))
compareML(durC_m, durC_m2)
```
Sign. improvement.


Add IPI as fixed effect.
```{r}
summary(durC_m3 <- bam(durationC ~ whichC*time.lag.norm + s(subj,bs='re') + s(subj, whichC, bs="re"), data = df.nospr, method = "ML"))
compareML(durC_m2, durC_m3)
```
Sign. improvement.

Add interaction with cluster.
```{r}
summary(durC_m4 <- bam(durationC ~ whichC*time.lag.norm + cluster + s(subj,bs='re') + s(subj, whichC, bs="re"), data = df.nospr, method = "ML"))
compareML(durC_m3, durC_m4)
```
No improvement.

Add gender as fixed effect.
```{r}
summary(durC_m5 <- bam(durationC ~ whichC*cluster + gender + s(time.lag.norm) + s(subj,bs='re') + s(subj, whichC, bs="re"), data = df.nospr, method = "ML"))
compareML(durC_m4, durC_m5)
```
No improvement.

Add speech rate as smooth
```{r}
summary(durC_m6 <- bam(durationC ~ whichC*cluster + s(duration.tt) + s(time.lag.norm) + s(subj,bs='re') + s(subj, whichC, bs="re"), data = df.nospr, method = "ML"))
```
No improvement.


Add recording.no as smooth
```{r}
summary(durC_m7 <- bam(durationC ~ whichC*time.lag.norm + s(recording.no) + s(subj,bs='re') + s(subj, whichC, bs="re"), data = df.nospr, method = "ML")) # p = 1?
```
No. improvement.

Add dialect as fixed effect.
```{r}
summary(durC_m7 <- bam(durationC ~ whichC*time.lag.norm + s(recording.no) + s(subj,bs='re') + s(subj, whichC, bs="re"), data = df.nospr, method = "ML")) # p = 1?
compareML(IPI_Cm5, IPI_Cm9)
```
No improvement.

Add age as smooth.
```{r}
summary(IPI_Cm10 <- bam(time.lag.norm ~ age + cluster +  s(durationC) + s(subj,bs='re') + s(subj, whichC, bs="re"), data = df.nospr, method = "ML"))
compareML(IPI_Cm5, IPI_Cm10)
```
No improvement.

Add random slopes cluster.
```{r}
summary(IPI_Cm5REML <- bam(time.lag.norm ~   cluster +  s(durationC) + s(subj,bs='re') + s(subj, whichC, bs="re"), data = df.nospr, method = "REML"))
summary(IPI_Cm11REML <- bam(time.lag.norm ~  cluster +  s(durationC) + s(subj,bs='re') + s(subj, whichC, bs="re") + s(subj, cluster, bs="re"), data = df.nospr, method = "REML"))
```
No improvement.

Add whichC as fixed factor:
```{r}
summary(IPI_Cm12 <- bam(time.lag.norm ~ cluster + whichC + s(durationC) + s(subj,bs='re') + s(subj, whichC, bs="re"), data = df.nospr, method = "ML"))
compareML(IPI_Cm5, IPI_Cm12)
```


Check for group once more (interaction with cluster)
```{r}
summary(IPI_Cm13 <- bam(time.lag.norm ~ group*cluster + whichC + s(durationC) + s(subj,bs='re') + s(subj, whichC, bs="re"), data = df.nospr, method = "ML"))
compareML(IPI_Cm13, IPI_Cm12)
```
Sign. improv.


Plot final model
```{r}
summary(IPI_Cm13REML <- bam(time.lag.norm ~ group*cluster + whichC + s(durationC) + s(subj,bs='re') + s(subj, whichC, bs="re"), data = df.nospr, method = "REML"))

gam.check(IPI_Cm13REML) #looks odd

summary(IPI_Cm13REML_scat <- bam(time.lag.norm ~ group*cluster + whichC + s(durationC) + s(subj,bs='re') + s(subj, whichC, bs="re"), data = df.nospr, method = "fREML", family = "scat", discrete = T))

gam.check(IPI_Cm13REML_scat) #looks a bit better

plot_smooth(IPI_Cm13REML, view = "durationC", plot_all=c("whichC", "group"), xlab = "Normalized C duration in /sp/", ylab = "Normalized inter plateau interval", cond=list(cluster = "sp"), ylim=c(-0.4,0.8))

plot_smooth(IPI_Cm13REML, view = "durationC", plot_all=c("whichC", "group"), xlab = "Normalized C duration in /sm/", ylab = "Normalized inter plateau interval", cond=list(cluster = "sm"), ylim=c(-0.4,0.8))

plot_smooth(IPI_Cm13REML, view = "durationC", plot_all=c("whichC", "group"), xlab = "Normalized C duration in /sx/", ylab = "Normalized inter plateau interval", cond=list(cluster = "sx"), ylim=c(-0.8,2))
```

