---
title: "C_center"
output:
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/45598770/Documents/analysis")
getwd()
```



```{r pressure, echo=FALSE}
require("dplyr")
require("tidyr")
require("ggplot2")
require("stringr")
```

## Read data

```{r read}
df <- read.csv("c-center_24Jun2020.csv", header = TRUE, sep = ",")
df_meta <- read.csv("metadata.csv", header = TRUE, sep = ",")
df <-  merge(df, df_meta, by="subj")
rm(df_meta)
```

```{r}
df_speech.rate <- read.csv("c-center_speech_rate.csv", header = TRUE, sep = ",")

# extract info from filename
df_speech.rate <- df_speech.rate %>%
    separate(fname, c("project", "prompt", "block", "remove", "remove2", "trialno", "repetition"), sep = "_")

# merge trialnumber and repetition
df_speech.rate <- df_speech.rate %>% 
  mutate(trial=str_c(trialno,repetition)) 

onset.interval <- df_speech.rate %>%
  group_by(subj) %>%
  group_by(trial) %>%
  filter(seg == "t1") %>%
  rename(utt.onset = t5)  %>%
  select(subj, trial, utt.onset)

offset.interval <- df_speech.rate %>%
  group_by(subj) %>%
  group_by(trial) %>%
  filter(seg == "t2") %>%
  rename(utt.offset = t3)  %>%
  select(subj, trial, utt.offset)

df_speech.rate <- merge(onset.interval, offset.interval, by=c("subj", "trial"))
rm(onset.interval, offset.interval)
```


## Preprocess data

```{r preprocess}
# extract info from filename
df<- df %>%
    separate(fname, c("project", "prompt", "block", "remove", "remove2", "trialno", "repetition"), sep = "_")

# merge trialnumber and repetition, add on- and offset of complete utterance
df <- df %>% 
  mutate(trial=str_c(trialno,repetition)) %>% 
  left_join(df_speech.rate, by = c("subj","trial"))

# create new variable recording.no to indicate the recording number (first recording, second recording etc..)
trialno <- df %>% 
group_by(subj) %>% 
distinct(trial) %>% 
arrange(trial, by_group = TRUE) %>%
mutate(recording.no = 1:n()) %>% 
select(recording.no, everything()) %>% 
arrange(subj, recording.no)

# bind trialno with df
df <- df %>% 
  inner_join(trialno, by = c("subj", "trial")) %>% 
  select(recording.no, everything()) %>% 
  arrange(subj, recording.no)

rm(trialno)

# create group variable
df$group <- ifelse((str_detect(df$subj, "CTRL")), "CTRL", "PD")
df$group <-  as.factor(df$group)

# drop unwanted columns
drop <- c("remove","remove2", "X")
df = df[!(names(df) %in% drop)]

# create condition based on prompt
df$condition <- ifelse(df$prompt == "oma spat" | df$prompt == "opa schat" | df$prompt == "opa Smat" | df$prompt == "oma sprak", "complex", "simple")

# create clustercolumn
df$cluster[df$prompt == "oma spat" | df$prompt == "oma's pad"] <- "sp" 
df$cluster[df$prompt == "opa schat" | df$prompt == "opa's gat"] <- "sx" 
df$cluster[df$prompt == "opa Smat" | df$prompt == "opa's mat"] <- "sm" 
df$cluster[df$prompt == "oma sprak" | df$prompt == "oma's prak"] <- "spr" 
df$cluster <- as.factor(df$cluster)
levels(df$cluster)



# remove faulty segmentation, based on outlier column
df <- df[is.na(df$outl),]

# check
table(df$prompt, df$condition)

table(df$seg, df$traj)

# remove unreliable sx segmentation of PD25
df <- df[!(df$subj == "PD25" & df$cluster == "sx"),]

```

# Speech rates

# Durations of rightmost consonant
```{r}
# collect duration /p/
duration.p <- df %>%
  filter(prompt == "oma spat" | prompt == "oma's pad" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "p") %>%
  mutate(duration = t5-t3)

ggplot(duration.p, aes(x=group, y=duration, fill=condition)) + geom_violin() + labs(title = "Duration /p/ in /sp/", y = "Duration")

# collect duration /m/
duration.m <- df %>%
  filter(prompt == "opa Smat" | prompt == "opa's mat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "m") %>%
  mutate(duration = t5-t3)

ggplot(duration.m, aes(x=group, y=duration, fill=condition)) + geom_violin() + labs(title = "Duration /m/ in /sm/", y = "Duration")

# collect duration /x/
duration.x <- df %>%
  filter(prompt == "opa schat" | prompt == "opa's gat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "X") %>%
  mutate(duration = t5-t3)

ggplot(duration.x, aes(x=group, y=duration, fill=condition)) + geom_violin() + labs(title = "Duration /x/ in /sx/", y = "Duration")

# collect duration /p/
duration.p.spr <- df %>%
  filter(prompt == "oma sprak" | prompt == "oma's prak" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "p") %>%
  mutate(duration = t5-t3)

ggplot(duration.p.spr, aes(x=group, y=duration, fill=condition)) + geom_violin() + labs(title = "Duration /p/ in /spr/", y = "Duration")

```


# Rightedge
Computing durations between the offset of the vowel's preceding consonant and onset of the vowel's following consonant (i.e. the anchor)

/sp/ environment
```{r sp}
# collect onset times onset /p/ (start of word)
onset.p <- df %>%
  filter(prompt == "oma spat" | prompt == "oma's pad" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "p") %>%
  select(subj, recording.no, prompt, block, condition, seg, group, cluster, utt.onset, utt.offset,  t3) %>%
  rename(t3.p = t3)

# collect onset times coda /t/ (end of word)
onset.t <- df %>%
  filter(prompt == "oma spat" | prompt == "oma's pad" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "t") %>%
  select(subj, recording.no, prompt, block, condition, seg, group, cluster, utt.onset, utt.offset,  t3) %>%
  rename(t3.t = t3)

# collect offset times /p/ (gesture of interest), merge dataframes and time normalize (0,1) the offset of /p/
sp <- df %>%
  filter(prompt == "oma spat" | prompt == "oma's pad" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "p") %>%
  select(subj, recording.no, prompt, block, condition, seg, group, cluster, utt.onset, utt.offset,  t5) %>%
  rename(t5.p = t5) %>%
  inner_join(onset.p, by = c("subj", "recording.no")) %>%
  inner_join(onset.t, by = c("subj", "recording.no")) %>%
  mutate(time.lag = t3.t-t5.p)  %>% 
  mutate(time.lag.norm =  1 - ((t5.p - utt.onset)/(utt.offset-utt.onset))) # time normalisation from 0 to 1 based on (x-min(x))/(max(x)-min(x))

sp <- sp %>%
    group_by(subj, prompt) %>%
    mutate(mean.lag.prompt = mean(time.lag, na.rm=TRUE)) %>%
    mutate(mean.lag.prompt.norm = mean(time.lag.norm, na.rm=TRUE)) 

  rm(onset.p, onset.t)
```

/sm/ environment
```{r sm}
# collect onset times onset /m/ (start of word)
onset.m <- df %>%
  filter(prompt == "opa Smat" | prompt == "opa's mat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "m") %>%
  select(subj, recording.no, prompt, block, condition, seg, group, cluster, utt.onset, utt.offset,  t3) %>%
  rename(t3.m = t3)

# collect onset times coda /t/ (end of word)
onset.t <- df %>%
  filter(prompt == "opa Smat" | prompt == "opa's mat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "t") %>%
  select(subj, recording.no, prompt, block, condition, seg, group, cluster, utt.onset, utt.offset,  t3) %>%
  rename(t3.t = t3)

# collect offset times /m/ (gesture of interest), merge dataframes and time normalize (0,1) the offset of /m/
sm <- df %>%
  filter(prompt == "opa Smat" | prompt == "opa's mat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "m") %>%
  select(subj, recording.no, prompt, block, condition, seg, group, cluster, utt.onset, utt.offset,  t5) %>%
  rename(t5.m = t5) %>%
  inner_join(onset.m, by = c("subj", "recording.no")) %>%
  inner_join(onset.t, by = c("subj", "recording.no")) %>%
  mutate(time.lag = t3.t-t5.m)  %>% 
  mutate(time.lag.norm = 1 - ((t5.m - utt.onset)/(utt.offset-utt.onset)))

sm <- sm %>%
    group_by(subj, prompt) %>%
    mutate(mean.lag.prompt = mean(time.lag, na.rm=TRUE)) %>%
    mutate(mean.lag.prompt.norm = mean(time.lag.norm, na.rm=TRUE))


  rm(onset.m, onset.t)
```

/sx/ environment
```{r sx}
# collect onset times onset /x/ (start of word)
onset.x <- df %>%
  filter(prompt == "opa schat" | prompt == "opa's gat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "X") %>%
  select(subj, recording.no, prompt, block, condition, seg, group, utt.onset, utt.offset, cluster,  t3) %>%
  rename(t3.x = t3)

# collect onset times coda /t/ (end of word)
onset.t <- df %>%
  filter(prompt == "opa schat" | prompt == "opa's gat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "t") %>%
  select(subj, recording.no, prompt, block, condition, seg, group, utt.onset, utt.offset, cluster,  t3) %>%
  rename(t3.t = t3)

# collect offset times /x/ (gesture of interest), merge dataframes and time normalize (0,1) the offset of /x/
sx <- df %>%
  filter(prompt == "opa schat" | prompt == "opa's gat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "X") %>%
  select(subj, recording.no, prompt, block, condition, seg, group, utt.onset, utt.offset, cluster,  t5) %>%
  rename(t5.x = t5) %>%
  inner_join(onset.x, by = c("subj", "recording.no")) %>%
  inner_join(onset.t, by = c("subj", "recording.no")) %>%
  mutate(time.lag = t3.t-t5.x)  %>% 
  mutate(time.lag.norm = 1 - ((t5.x - utt.onset)/(utt.offset-utt.onset)))

sx <- sx %>%
    group_by(subj, prompt) %>%
    mutate(mean.lag.prompt = mean(time.lag, na.rm=TRUE)) %>%
    mutate(mean.lag.prompt.norm = mean(time.lag.norm, na.rm=TRUE))


  rm(onset.x, onset.t)
```

/spr/ environment
```{r spr}
# collect onset times onset /s/ (start of word)
onset.p <- df %>%
  filter(prompt == "oma sprak" | prompt == "oma's prak" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "p") %>%
  select(subj, recording.no, prompt, block, condition, seg, group, cluster, utt.onset, utt.offset,  t3) %>%
  rename(t3.p = t3)

# collect onset times coda /k/ (end of word)
onset.k <- df %>%
  filter(prompt == "oma sprak" | prompt == "oma's prak" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "k") %>%
  select(subj, recording.no, prompt, block, condition, seg, group, cluster, utt.onset, utt.offset,  t3) %>%
  rename(t3.k = t3)

# collect offset times /r/ (gesture of interest), merge dataframes and time normalize (0,1) the offset of /r/
spr <- df %>%
  filter(prompt == "oma sprak" | prompt == "oma's prak" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "r") %>%
  select(subj, recording.no, prompt, block, condition, seg, group, cluster, utt.onset, utt.offset,  t5) %>%
  rename(t5.r = t5) %>%
  inner_join(onset.p, by = c("subj", "recording.no")) %>%
  inner_join(onset.k, by = c("subj", "recording.no")) %>%
  mutate(time.lag = t3.k-t5.r)  %>% 
  mutate(time.lag.norm = 1 - ((t5.r - utt.onset)/(utt.offset-utt.onset)))

spr <- spr %>%
    group_by(subj, prompt) %>%
    mutate(mean.lag.prompt = mean(time.lag, na.rm=TRUE)) %>%
    mutate(mean.lag.prompt.norm = mean(time.lag.norm, na.rm=TRUE))


  rm(onset.p, onset.k)
```

Merge.
```{r merge}
df.rightedge <- rbind(sm, sp, sx, spr)
# df.rightedge <- unique(df.rightedge[,c("subj", "group", "condition","cluster","mean.lag.prompt.norm")])
rm(sm, sp, sx, spr)

```


Plots
```{r plots}
#right edge shift
ggplot(df.rightedge, aes(x=group, y=mean.lag.prompt, fill=condition)) + geom_violin() + labs(title = "Overall", y = "Distance to anchor")

dodge <- position_dodge(width = 0.9)

#/sp/
ggplot(df.rightedge[df.rightedge$cluster == "sp",], aes(x=group, y=mean.lag.prompt, fill=condition)) + geom_violin() + labs(title = "sp", y = "Lag") + geom_boxplot(width=.2, position = dodge)

#/sm/
ggplot(df.rightedge[df.rightedge$cluster == "sm",], aes(x=group, y=mean.lag.prompt, fill=condition)) + geom_violin() + labs(title = "sm", y = "Lag") + geom_boxplot(width=.2, position = dodge)


#/sX/
ggplot(df.rightedge[df.rightedge$cluster == "sx",], aes(x=group, y=mean.lag.prompt, fill=condition)) + geom_violin() + labs(title = "sx", y = "Lag") + geom_boxplot(width=.2, position = dodge)

#/spr/
ggplot(df.rightedge[df.rightedge$cluster == "spr",], aes(x=group, y=mean.lag.prompt, fill=condition)) + geom_violin() + labs(title = "spr", y = "Lag") + geom_boxplot(width=.2, position = dodge)


ggplot(df.rightedge[df.rightedge$cluster == "sm",], aes(x=subj, y=time.lag.norm, fill=condition)) +
geom_bar(stat="identity", position=position_dodge())

ggplot(df.rightedge[df.rightedge$cluster == "sp",], aes(x=subj, y=time.lag.norm, fill=condition)) +
geom_bar(stat="identity", position=position_dodge())

ggplot(df.rightedge[df.rightedge$cluster == "sx",], aes(x=subj, y=time.lag.norm, fill=condition)) +
geom_bar(stat="identity", position=position_dodge())

ggplot(df.rightedge[df.rightedge$cluster == "spr",], aes(x=subj, y=time.lag.norm, fill=condition)) +
geom_bar(stat="identity", position=position_dodge())
```
# Pairwise comparisons

/sp/
```{r}
sp.simple <- sp %>%
    ungroup()  %>%
    filter(condition == "simple") %>%
    rename(time.lag.simple = time.lag) %>%
    select(subj, time.lag.simple) %>%
    drop_na() %>%
    group_by(subj) %>%
    mutate(rep = 1:n())


sp.complex <- sp %>%
    ungroup()  %>%
    filter(condition == "complex") %>%
    rename(time.lag.complex = time.lag)  %>%
    select(subj, group, cluster, time.lag.complex) %>%
    drop_na() %>%    
    group_by(subj) %>%
    mutate(rep = 1:n())

sp.pairwise <- drop_na(merge(sp.complex, sp.simple, by = c("subj", "rep"), all = TRUE))
sp.pairwise$time.lag <- sp.pairwise$time.lag.simple - sp.pairwise$time.lag.complex
rm(sp.simple, sp.complex)
```

/sm/
```{r}
sm.simple <- sm %>%
    ungroup()  %>%
    filter(condition == "simple") %>%
    rename(time.lag.simple = time.lag) %>%
    select(subj, time.lag.simple) %>%
    drop_na() %>%
    group_by(subj) %>%
    mutate(rep = 1:n())


sm.complex <- sm %>%
    ungroup()  %>%
    filter(condition == "complex") %>%
    rename(time.lag.complex = time.lag)  %>%
    select(subj, group, cluster, time.lag.complex) %>%
    drop_na() %>%    
    group_by(subj) %>%
    mutate(rep = 1:n())

sm.pairwise <- drop_na(merge(sm.complex, sm.simple, by = c("subj", "rep"), all = TRUE))
sm.pairwise$time.lag <- sm.pairwise$time.lag.simple - sm.pairwise$time.lag.complex
rm(sm.simple, sm.complex)
```

/sx/
```{r}
sx.simple <- sx %>%
    ungroup()  %>%
    filter(condition == "simple") %>%
    rename(time.lag.simple = time.lag) %>%
    select(subj, time.lag.simple) %>%
    drop_na() %>%
    group_by(subj) %>%
    mutate(rep = 1:n())


sx.complex <- sx %>%
    ungroup()  %>%
    filter(condition == "complex") %>%
    rename(time.lag.complex = time.lag)  %>%
    select(subj, group, cluster, time.lag.complex) %>%
    drop_na() %>%    
    group_by(subj) %>%
    mutate(rep = 1:n())

sx.pairwise <- drop_na(merge(sx.complex, sx.simple, by = c("subj", "rep"), all = TRUE))
sx.pairwise$time.lag <- sx.pairwise$time.lag.simple - sx.pairwise$time.lag.complex
rm(sx.simple, sx.complex)
```

/spr/
```{r}
spr.simple <- spr %>%
    ungroup()  %>%
    filter(condition == "simple") %>%
    rename(time.lag.simple = time.lag) %>%
    select(subj, time.lag.simple) %>%
    drop_na() %>%
    group_by(subj) %>%
    mutate(rep = 1:n())


spr.complex <- spr %>%
    ungroup()  %>%
    filter(condition == "complex") %>%
    rename(time.lag.complex = time.lag)  %>%
    select(subj, group, cluster, time.lag.complex) %>%
    drop_na() %>%    
    group_by(subj) %>%
    mutate(rep = 1:n())

spr.pairwise <- drop_na(merge(spr.complex, spr.simple, by = c("subj", "rep"), all = TRUE))
spr.pairwise$time.lag <- spr.pairwise$time.lag.simple - spr.pairwise$time.lag.complex
rm(spr.simple, spr.complex)
```

Bind together
```{r}
df.pairwise <- rbind(sp.pairwise, sm.pairwise, sx.pairwise, spr.pairwise)
rm(sp.pairwise, sm.pairwise, sx.pairwise, spr.pairwise)
```


# Leftedge

```{r leftedge shift}
# collect onset times onset /s/ (start of word)
onset.p <- df %>%
  filter(prompt == "oma's prak" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "p") %>%
  select(subj, recording.no, prompt, block, condition, seg, group, cluster,  t3) %>%
  rename(t3.p = t3)

onset.k <- df %>%
  filter(prompt == "oma's prak" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "k") %>%
  select(subj, recording.no, prompt, block, condition, seg, group, cluster,  t3) %>%
  rename(t3.k = t3)


pr <- df %>%
  filter(prompt == "oma's prak" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "p") %>%
  select(subj, recording.no, prompt, block, condition, seg, group, cluster,  t5) %>%
  rename(t5.p = t5) %>%
  inner_join(onset.p, by = c("subj", "recording.no", "prompt", "block", "condition", "group", "cluster")) %>%
  inner_join(onset.k, by = c("subj", "recording.no", "prompt", "block", "condition", "group", "cluster")) %>%
  mutate(time.lag.norm = (t5.p - t3.p)/(t3.k-t3.p)) # based on (x-min(x))/(max(x)-min(x))


# collect onset times onset /s/ (start of word)
onset.p <- df %>%
  filter(prompt == "oma's pad" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "p") %>%
  select(subj, recording.no, prompt, block, condition, seg, group, cluster,  t3) %>%
  rename(t3.p = t3)

onset.t <- df %>%
  filter(prompt == "oma's pad" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "t") %>%
  select(subj, recording.no, prompt, block, condition, seg, group, cluster,  t3) %>%
  rename(t3.t = t3)


p <- df %>%
  filter(prompt == "oma's pad" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "p") %>%
  select(subj, recording.no, prompt, block, condition, seg, group, cluster,  t5) %>%
  rename(t5.p = t5) %>%
  inner_join(onset.p, by = c("subj", "recording.no", "prompt", "block", "condition", "group", "cluster")) %>%
  inner_join(onset.t, by = c("subj", "recording.no", "prompt", "block", "condition", "group", "cluster")) %>%
  mutate(time.lag.norm = (t5.p - t3.p)/(t3.t-t3.p)) # based on (x-min(x))/(max(x)-min(x))


df.leftedge <- rbind(pr, p)
df.leftedge$condition <- ifelse((str_detect(df.leftedge$prompt, "oma's pad")), "simple", "complex")
```


# CC clusters

/sp/ CC lag
```{r sp CC lag}
# collect onset times /s/ (start of interval)
onset.s <- df %>%
  filter(prompt == "oma's pad" | prompt == "oma spat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "s") %>%
  select(subj, recording.no, prompt, block, condition, group, cluster,  t3) %>%
  rename(t3.s = t3)

# collect offset times /p/ (end of interval)
offset.p <- df %>%
  filter(prompt == "oma's pad" | prompt == "oma spat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "p") %>%
  select(subj, recording.no, prompt, block, condition, group, cluster,  t5) %>%
  rename(t5.p = t5)

# collect normalized offset times /s/
offset.s.norm <- df %>%
  filter(prompt == "oma's pad" | prompt == "oma spat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "s") %>%
  select(subj, recording.no, prompt, block, condition, group, cluster,  t5) %>%
  rename(t5.s = t5)  %>%
  inner_join(onset.s, by = c("subj", "recording.no")) %>%
  inner_join(offset.p, by = c("subj", "recording.no")) %>%
  mutate(offset.s.norm = (t5.s - t3.s)/(t5.p-t3.s)) # time normalisation from 0 to 1 based on (x-min(x))/(max(x)-min(x))
  
  # collect normalized offset times /s/
sp.CC.norm <- df %>%
  filter(prompt == "oma's pad" | prompt == "oma spat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "p") %>%
  select(subj, recording.no, prompt, block, condition, group, cluster,  t3) %>%
  rename(t3.p = t3)  %>%
  inner_join(onset.s, by = c("subj", "recording.no", "prompt", "block", "condition", "group", "cluster")) %>%
  inner_join(offset.p, by = c("subj", "recording.no", "prompt", "block", "condition", "group", "cluster")) %>%
  mutate(onset.p.norm = (t3.p - t3.s)/(t5.p-t3.s)) %>% # time normalisation from 0 to 1 based on (x-min(x))/(max(x)-min(x))
  inner_join(offset.s.norm, by = c("subj", "recording.no", "prompt", "block", "condition", "group", "cluster")) %>%
  mutate(CC.lag.norm = onset.p.norm - offset.s.norm)

rm(onset.s, offset.p, offset.s.norm)
```

/sm/ CC lag
```{r sm CC lag}
# collect onset times /s/ (start of interval)
onset.s <- df %>%
  filter(prompt == "opa Smat" | prompt == "opa's mat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "s") %>%
  select(subj, recording.no, prompt, block, condition, group, cluster,  t3) %>%
  rename(t3.s = t3)

# collect offset times /m/ (end of interval)
offset.m <- df %>%
  filter(prompt == "opa Smat" | prompt == "opa's mat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "m") %>%
  select(subj, recording.no, prompt, block, condition, group, cluster,  t5) %>%
  rename(t5.m = t5)

# collect normalized offset times /s/
offset.s.norm <- df %>%
  filter(prompt == "opa Smat" | prompt == "opa's mat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "s") %>%
  select(subj, recording.no, prompt, block, condition, group, cluster,  t5) %>%
  rename(t5.s = t5)  %>%
  inner_join(onset.s, by = c("subj","recording.no", "block", "condition", "group", "cluster")) %>%
  inner_join(offset.m, by = c("subj", "recording.no", "block", "condition", "group", "cluster")) %>%
  mutate(offset.s.norm = (t5.s - t3.s)/(t5.m-t3.s)) # time normalisation from 0 to 1 based on (x-min(x))/(max(x)-min(x))
  
  # collect normalized offset times /m/
sm.CC.norm <- df %>%
  filter(prompt == "opa Smat" | prompt == "opa's mat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "m") %>%
  select(subj, recording.no, prompt, block, condition, group, cluster,  t3) %>%
  rename(t3.m = t3)  %>%
  inner_join(onset.s, by = c("subj", "recording.no", "prompt", "block", "condition", "group", "cluster")) %>%
  inner_join(offset.m, by = c("subj", "recording.no", "prompt", "block", "condition", "group", "cluster")) %>%
  mutate(onset.m.norm = (t3.m - t3.s)/(t5.m-t3.s)) %>% # time normalisation from 0 to 1 based on (x-min(x))/(max(x)-min(x))
  inner_join(offset.s.norm, by = c("subj", "recording.no", "prompt", "block", "condition", "group", "cluster")) %>%
  mutate(CC.lag.norm = onset.m.norm - offset.s.norm)

rm(onset.s, offset.m, offset.s.norm)
```


/sx/ CC lag
```{r}
# collect onset times /s/ (start of interval)
onset.s <- df %>%
  filter(prompt == "opa schat" | prompt == "opa's gat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "s") %>%
  select(subj, recording.no, prompt, block, condition, group, cluster,  t3) %>%
  rename(t3.s = t3)

# collect offset times /x/ (end of interval)
offset.x <- df %>%
  filter(prompt == "opa schat" | prompt == "opa's gat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "X") %>%
  select(subj, recording.no, prompt, block, condition, group, cluster,  t5) %>%
  rename(t5.x = t5)

# collect normalized offset times /s/
offset.s.norm <- df %>%
  filter(prompt == "opa schat" | prompt == "opa's gat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "s") %>%
  select(subj, recording.no, prompt, block, condition, group, cluster,  t5) %>%
  rename(t5.s = t5)  %>%
  inner_join(onset.s, by = c("subj", "recording.no", "block", "condition", "group", "cluster")) %>%
  inner_join(offset.x, by = c("subj", "recording.no", "block", "condition", "group", "cluster")) %>%
  mutate(offset.s.norm = (t5.s - t3.s)/(t5.x-t3.s)) # time normalisation from 0 to 1 based on (x-min(x))/(max(x)-min(x))
  
  # collect normalized offset times /x/
sx.CC.norm <- df %>%
  filter(prompt == "opa schat" | prompt == "opa's gat" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "X") %>%
  select(subj, recording.no, prompt, block, condition, group, cluster,  t3) %>%
  rename(t3.x = t3)  %>%
  inner_join(onset.s, by = c("subj", "recording.no", "prompt", "block", "condition", "group", "cluster")) %>%
  inner_join(offset.x, by = c("subj", "recording.no", "prompt", "block", "condition", "group", "cluster")) %>%
  mutate(onset.x.norm = (t3.x - t3.s)/(t5.x-t3.s)) %>% # time normalisation from 0 to 1 based on (x-min(x))/(max(x)-min(x))
  inner_join(offset.s.norm, by = c("subj", "recording.no", "prompt", "block", "condition", "group", "cluster")) %>%
  mutate(CC.lag.norm = onset.x.norm - offset.s.norm)

rm(onset.s, offset.x, offset.s.norm)
```


/spr/ CC lag
```{r}
# collect onset times /s/ (start of interval)
onset.s <- df %>%
  filter(prompt == "oma sprak" | prompt == "oma's prak" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "s") %>%
  select(subj, recording.no, prompt, block, condition, group, cluster,  t3) %>%
  rename(t3.s = t3)

# collect offset times /r/ (end of interval)
offset.r <- df %>%
  filter(prompt == "oma sprak" | prompt == "oma's prak" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "r") %>%
  select(subj, recording.no, prompt, block, condition, group, cluster,  t5) %>%
  rename(t5.r = t5)

# collect normalized offset times /s/
offset.s.norm <- df %>%
  filter(prompt == "oma sprak" | prompt == "oma's prak" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "s") %>%
  select(subj, recording.no, prompt, block, condition, group, cluster,  t5) %>%
  rename(t5.s = t5)  %>%
  inner_join(onset.s, by = c("subj", "recording.no", "block", "condition", "group", "cluster")) %>%
  inner_join(offset.r, by = c("subj", "recording.no", "block", "condition", "group", "cluster")) %>%
  mutate(offset.s.norm = (t5.s - t3.s)/(t5.r-t3.s)) # time normalisation from 0 to 1 based on (r-min(r))/(mar(r)-min(r))
  
  # collect normalized offset times /r/
spr.CC.norm <- df %>%
  filter(prompt == "oma sprak" | prompt == "oma's prak" ) %>%
  group_by(subj) %>%
  group_by(recording.no) %>%
  filter(seg == "r") %>%
  select(subj, recording.no, prompt, block, condition, group, cluster,  t3) %>%
  rename(t3.r = t3)  %>%
  inner_join(onset.s, by = c("subj", "recording.no", "prompt", "block", "condition", "group", "cluster")) %>%
  inner_join(offset.r, by = c("subj", "recording.no", "prompt", "block", "condition", "group", "cluster")) %>%
  mutate(onset.r.norm = (t3.r - t3.s)/(t5.r-t3.s)) %>% # time normalisation from 0 to 1 based on (r-min(r))/(mar(r)-min(r))
  inner_join(offset.s.norm, by = c("subj", "recording.no", "prompt", "block", "condition", "group", "cluster")) %>%
  mutate(CC.lag.norm = onset.r.norm - offset.s.norm)

rm(onset.s, offset.r, offset.s.norm)
```

Merge.
```{r merge}
df.CC.lag<- rbind(sm.CC.norm, sp.CC.norm, sx.CC.norm, spr.CC.norm)
rm(sm.CC.norm, sp.CC.norm, sx.CC.norm, spr.CC.norm)
# df.rightedge <- unique(df.rightedge[,c("subj", "group", "condition","cluster","mean.lag.prompt.norm")])
```


## Some additional plots

```{r plots}
#right edge shift
ggplot(df.rightedge, aes(x=group, y=time.lag.norm, fill=condition)) + geom_violin() + labs(title = "Overall", y = "Distance to anchor")

dodge <- position_dodge(width = 0.9)

#left edge shift
ggplot(df.leftedge, aes(x=group, y=time.lag.norm, fill=prompt)) + geom_violin() + labs(title = "Overall", y = "Distance to anchor (ms)") + geom_boxplot(width=.2, position = dodge) + ylim(0,400)

#CC lag /sp/
ggplot(df.CC.lag[df.CC.lag$cluster == "sp",], aes(x=group, y=CC.lag.norm, fill=condition)) + geom_violin() + labs(title = "Overall", y = "Lag between /s/ and /p/ (ms)") + geom_boxplot(width=.2, position = dodge)

#CC lag /sm/
ggplot(df.CC.lag[df.CC.lag$cluster == "sm",], aes(x=group, y=CC.lag.norm, fill=condition)) + geom_violin() + labs(title = "Overall", y = "Lag between /s/ and /m/ (ms)") + geom_boxplot(width=.2, position = dodge)


#CC lag /sX/
ggplot(df.CC.lag[df.CC.lag$cluster == "sx",], aes(x=group, y=CC.lag.norm, fill=condition)) + geom_violin() + labs(title = "Overall", y = "Lag between /s/ and /x/ (ms)") + geom_boxplot(width=.2, position = dodge)


#CC lag /spr/
ggplot(df.CC.lag[df.CC.lag$cluster == "spr",], aes(x=group, y=CC.lag.norm, fill=condition)) + geom_violin() + labs(title = "Overall", y = "Lag between /s/ and /p/ (ms)")
```

```{r save}
#add meta_data
setwd("/Users/45598770/Documents/analysis")
df_meta <- read.csv("metadata.csv", header = TRUE, sep = ",")
df.leftedge <-  merge(df.leftedge, df_meta, by="subj")
df.rightedge <-  merge(df.rightedge, df_meta, by="subj")
df.CC.lag <-  merge(df.CC.lag, df_meta, by="subj")
df.pairwise <-  merge(df.pairwise, df_meta, by="subj")
rm(df_meta)


# Set variables that consist most datapoints as reference dummy variables
df.leftedge$gender <- relevel(df.leftedge$gender, ref = "M")
df.leftedge$group <- relevel(df.leftedge$group, ref = "CTRL")
df.leftedge$dialect <- relevel(df.leftedge$dialect, ref = "Rest")

# Set variables that consist most datapoints as reference dummy variables
df.rightedge$gender <- relevel(df.rightedge$gender, ref = "M")
df.rightedge$group <- relevel(df.rightedge$group, ref = "CTRL")
df.rightedge$dialect <- relevel(df.rightedge$dialect, ref = "Rest")

# Set variables that consist most datapoints as reference dummy variables
df.CC.lag$gender <- relevel(df.CC.lag$gender, ref = "M")
df.CC.lag$group <- relevel(df.CC.lag$group, ref = "CTRL")
df.CC.lag$dialect <- relevel(df.CC.lag$dialect, ref = "Rest")

save(df.pairwise, file="modelling_data_pairwise.Rda")
save(df.leftedge, file="modelling_data_leftedge.Rda")
save(df.rightedge, file="modelling_data_rightedge.Rda")
save(df.CC.lag, file="modelling_data_CC.Rda")
```

