---
title: "C-center models"
output: 
  html_document: 
    number_sections: yes
    toc: yes
    toc_float: yes
    df_print: paged
---

# Load required packages
```{r}
library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)
library(lme4)
library(car)
library(visreg)
library(optimx)
library(lmerTest)
```
# Introduction 
From the c-center hypothesis it follows that onsets are coordinated inphase with the following vowel, meaning that they are initiated roughly around the same time. On the other hand, multiple consonants in an onset are also coordinated antiphase with each other (i.e. sequential initiation). To cope with these competing relations, the leftmost C gets shifted towards the left and the rightmost C gets shifted towards the right (with respect to the following vowel) in complex onsets. In this way, the average onset of the two C's is still initiated at the same time as the vowel, whereas they are also still coordinated in an antiphase relation with each other. Our hypothesis is that PD patients will show a stronger preference for the inphase relationship, coming at the cost of the antiphase relationship. In our data we've included three singular onsets /p/, /m/ and /x/ and four complex onsets: /sp/, /sm/, /sx/ and /spr/. The variable 'time.lag' is the time normalized (on a 0-1 scale) lag between consonant and vowel. 

## Read data
```{r save data}
load("../data/modelling_data_rightedge.Rda")
```

## Convert factors and subset data
```{r remove NA}
# convert to factors/numeric:
df.rightedge$condition <- as.factor(df.rightedge$condition)

# remove all the recordings without a speech rate measure:
df.rightedge <- df.rightedge %>% drop_na(duration.tt)

# center duration.tt, so it's easier to interpret.
df.rightedge$duration.tt.c <- scale(df.rightedge$duration.tt, scale = F)

# create df without spr:
df.rightedge.nospr <- df.rightedge[! df.rightedge$cluster == "spr", ]
df.rightedge.nospr <- droplevels(df.rightedge.nospr)
levels(df.rightedge.nospr$cluster)

# create df without sx:
#df.rightedge <- df.rightedge[! df.rightedge$cluster == "sx", ]
#df.rightedge <- droplevels(df.rightedge)
#levels(df.rightedge$cluster)
```


# LME models for rightward shift rightmost C
In a CCV complex onset the rightmost C in the onset should show a rightward shift towards the vowel in comparison to the C in a CV onset. We hypothesize that this shift may be less pronounced for the patient group, as they may show a stronger preference for inphase coordination. In this analysis we compare /sp/ to /p/, /sm/ to /m/, /sx/ to /x/ and /spr/ to /pr/. As a start, I've included the 'cluster' variable and the 'condition' (simple vs. complex onset) variable.

## Some plots
```{r}
#right edge shift
ggplot(df.rightedge, aes(x=group, y=time.lag, fill=condition)) + geom_violin() + labs(title = "Over all clusters", y = "Distance to anchor")
dodge <- position_dodge(width = 0.9)

ggplot(df.rightedge[df.rightedge$cluster == "sp",], aes(x=group, y=time.lag, fill=condition)) + geom_violin(trim=FALSE) + labs(title = "/sp/", y = "Distance to anchor") + geom_boxplot(width=.2, position = dodge) 

ggplot(df.rightedge[df.rightedge$cluster == "sm",], aes(x=group, y=time.lag, fill=condition)) + geom_violin(trim=FALSE) + labs(title = "/sm/", y = "Distance to anchor") + geom_boxplot(width=.2, position = dodge) 

ggplot(df.rightedge[df.rightedge$cluster == "spr",], aes(x=group, y=time.lag, fill=condition)) + geom_violin(trim=FALSE) + labs(title = "/spr/", y = "Distance to anchor") + geom_boxplot(width=.2, position = dodge) 
```

## Hypothesis testing

Fitting first model:
```{r}
summary(m <- lmer(time.lag ~ (1+condition + cluster|subj), data=df.rightedge, control = lmerControl(optimizer = "bobyqa")))$varcor
```

Adding cluster as fixed effect:
```{r}
summary(m1 <- lmer(time.lag ~ cluster + (1+condition + cluster|subj), data=df.rightedge, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m, m1)
```
Sign. improvement.

Adding condition as fixed effect:
```{r}
summary(m2 <- lmer(time.lag ~ condition + cluster + (1+condition + cluster|subj), data=df.rightedge, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m1, m2)
```
Sign. improvement.

Adding speech rate as fixed effect:
```{r}
summary(m3 <- lmer(time.lag ~ condition + cluster + duration.tt.c + (1+condition + cluster|subj), data=df.rightedge, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m2, m3)
```
Sign. improvement.

Adding interaction duration.tt:condition:
```{r}
summary(m4 <- lmer(time.lag ~ cluster + duration.tt.c*condition + (1+condition + cluster|subj), data=df.rightedge, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m3, m4)
```
No improvement.

Adding interaction duration.tt:cluster:
```{r}
summary(m5 <- lmer(time.lag ~ duration.tt.c*cluster + condition + (1+condition + cluster|subj), data=df.rightedge, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m3, m5)
```
No improvement.

Adding group as fixed effect:
```{r}
summary(m6 <- lmer(time.lag ~ condition + cluster + duration.tt.c + group + (1+condition + cluster|subj), data=df.rightedge, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m3, m6)
```
No improvement.

Adding condition:cluster as fixed effect.
```{r}
summary(m7 <- lmer(time.lag ~ condition*cluster + duration.tt.c + (1+condition + cluster|subj), data=df.rightedge, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m3, m7)
```
Sign. improvement.

Check whether we need to include condition:cluster in the random structure.
```{r}
summary(m8 <- lmer(time.lag ~ condition*cluster + duration.tt.c + (1+condition*cluster|subj), data=df.rightedge, control = lmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=2e5))))$coef
anova(m7, m8, refit = F)
```
Fails to converge and not a better fit. Continue with m7.

Adding group:condition as fixed effect (this is the main hypothesis).
```{r}
summary(m9 <- lmer(time.lag ~ group*condition + condition*cluster + duration.tt.c + (1+condition + cluster|subj), data=df.rightedge, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m7, m9) # is this a correct comparison?
```
No improvement.

Adding threeway interaction group:condition:cluster:
```{r}
summary(m10 <- lmer(time.lag ~ group*condition*cluster + duration.tt.c + (1+condition + cluster|subj), data=df.rightedge, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m7, m10)
```
No improvement.

Check whether we need random slopes for condition (by excluding condition from the random effect structure):
```{r}
summary(m11 <- lmer(time.lag ~ condition*cluster + duration.tt.c + (1+cluster|subj), data=df.rightedge, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m7, m11, refit=F)
```
We do need the random slope.

Check whether we really need random slopes for cluster (by excluding cluster from the random effect structure):
```{r}
summary(m12 <- lmer(time.lag ~ condition*cluster + duration.tt.c + (1+condition|subj), data=df.rightedge, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m7, m12, refit=F)
```
Again, we do need the random slope.


### Visualize final model:
```{r}
visreg(m5, "condition", by = "cluster", overlay = T, points = list(cex = 0.1), ylab = "time lag (ms)")
```


## Models without /spr/ cluster
Fitting first model:
```{r}
summary(m <- lmer(time.lag ~ (1+condition + cluster|subj), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$varcor
```

Adding cluster as fixed effect:
```{r}
summary(m1 <- lmer(time.lag ~ cluster + (1+condition + cluster|subj), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m, m1)
```
Sign. improvement.

Adding condition as fixed effect:
```{r}
summary(m2 <- lmer(time.lag ~ condition + cluster + (1+condition + cluster|subj), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m1, m2)
```
Sign. improvement.

Adding speech rate as fixed effect:
```{r}
summary(m3 <- lmer(time.lag ~ condition + cluster + duration.tt.c + (1+condition + cluster|subj), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m2, m3)
```
Sign. improvement.

Adding interaction condition:duration.tt as fixed effect:
```{r}
summary(m4 <- lmer(time.lag ~ condition*duration.tt.c + cluster + duration.tt.c + (1+condition + cluster|subj), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m3, m4)
```
No improvement.

Adding interaction cluster:duration.tt as fixed effect:
```{r}
summary(m5 <- lmer(time.lag ~ cluster*duration.tt.c + condition + (1+condition + cluster|subj), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m3, m5)
```
No improvement.

Adding condition:group interaction (main hypothesis): 
```{r}
summary(m6 <- lmer(time.lag ~ condition*group + cluster + duration.tt.c + (1+condition + cluster |subj), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m3, m6)
```
No improvement.

Adding duration:group interaction:
```{r}
summary(m7 <- lmer(time.lag ~ condition + cluster + duration.tt.c*group + (1+condition + cluster|subj), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m3, m7)
```
Sign. improvement.

Adding duration:group:condition interaction:
```{r}
summary(m8 <- lmer(time.lag ~ duration.tt.c*group*condition + cluster + (1+condition + cluster|subj), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$coef
```

Adding condition:cluster as fixed effect.
```{r}
summary(m9 <- lmer(time.lag ~ condition*cluster + duration.tt.c*group + (1+condition + cluster|subj), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m7, m9)
```
Sign. improvement.

Check whether we need to include condition:cluster in the random structure.
```{r}
summary(m10 <- lmer(time.lag ~ condition*cluster + duration.tt.c*group + (1+condition*cluster|subj), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m9, m10, refit = F)
```
Fails to converge. Continue with m10.

Check whether we need random slopes for condition:
```{r}
summary(m11 <- lmer(time.lag ~ condition*cluster + duration.tt.c*group + (1+cluster|subj), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m9, m11, refit=F)
```
We do need the random slope.

Check whether we really need random slopes for cluster:
```{r}
summary(m12 <- lmer(time.lag ~ condition*cluster + duration.tt.c*group + (1+condition|subj), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m9, m12, refit=F)
```
Again, we do need the random slope.


Visualize final model:
```{r}
visreg(m9, "cluster", by = "condition", overlay = T, points = list(cex = 0.1), ylab = "time lag (ms)")
```

## Exploratory analysis
Best model, including hypothesis:
```{r}
summary(m1 <- lmer(time.lag ~ condition*cluster + condition*group + duration.tt.c*group + (1+condition + cluster|subj), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$coef
```

Adding recording.no as fixed effect:
```{r}
summary(m2 <- lmer(time.lag ~ condition*cluster + condition*group + duration.tt.c*group + recording.no + (1+condition + cluster|subj), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m1,m2)
```
Sign. improvement.

Adding gender as fixed effect:
```{r}
summary(m3 <- lmer(time.lag ~ condition*cluster + condition*group + duration.tt.c*group + recording.no + gender + (1+condition + cluster|subj), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m2,m3)
```
Sign. improvement.

Adding dialect as fixed effect:
```{r}
summary(m4 <- lmer(time.lag ~ condition*cluster + condition*group + duration.tt.c*group + recording.no + gender + dialect + (1+condition + cluster|subj), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m3,m4)
```
No improvement. 

Adding age as fixed effect:
```{r}
summary(m5 <- lmer(time.lag ~ condition*cluster + condition*group + duration.tt.c*group + recording.no + gender + age + (1+condition + cluster|subj), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m3,m5)
```
No improvement.

Adding random intercepts for recording.no:
```{r}
summary(m6 <- lmer(time.lag ~ condition*cluster + condition*group + duration.tt.c*group + recording.no + gender + (1+condition + cluster|subj) + (1|recording.no), data=df.rightedge.nospr, control = lmerControl(optimizer = "bobyqa")))$coef
anova(m3,m6, refit = F)
```

### Visualize final model:
```{r}
visreg(m3, "cluster", by = "condition", overlay = T, points = list(cex = 0.1), ylab = "time lag (ms)")
```

## Checking assumptions final model

Check multicollinearity:
```{r}
#vif(m10) # OK, all lower than 5
```

Check autocorrelation:
```{r}
#acf(resid(m10)) # OK
```

Check normality residuals:
```{r}
#qnorm(resid(m10)) 
#qqline(resid(m10)) # doesn't look OK, short tailed?
```

Check hetereoscedasticity:
```{r}
#plot(fitted(m10), resid(m10)) # looks OK
```

## Models without controls

Best model, without control data and group variable.
```{r}
summary(m1 <- lmer(time.lag ~ condition*cluster + condition + duration.tt.c + recording.no + gender + (1+condition + cluster|subj), data=df.rightedge.nospr[df.rightedge.nospr$group == "PD",], control = lmerControl(optimizer = "bobyqa")))$coef
```

Include 'years since diagnosis' as fixed effect:
```{r}
summary(m2 <- lmer(time.lag ~ condition*cluster + condition + duration.tt.c + recording.no + gender + diagnosis + (1+condition + cluster|subj), data=df.rightedge.nospr[df.rightedge.nospr$group == "PD",], control = lmerControl(optimizer = "bobyqa")))$coef
```
No improvement.

Include 'tremor' as fixed effect:
```{r}
summary(m3 <- lmer(time.lag ~ condition*cluster + condition + duration.tt.c + recording.no + gender + tremor + (1+condition + cluster|subj), data=df.rightedge.nospr[df.rightedge.nospr$group == "PD",], control = lmerControl(optimizer = "bobyqa")))$coef
```
No improvement.

Include 'akinetic' as fixed effect:
```{r}
summary(m4 <- lmer(time.lag ~ condition*cluster + condition + duration.tt.c + recording.no + gender + akinetic + (1+condition + cluster|subj), data=df.rightedge.nospr[df.rightedge.nospr$group == "PD",], control = lmerControl(optimizer = "bobyqa")))$coef
anova(m1, m4)
```
Sign. improvement.

Include 'akinetic:condition' interaction as fixed effect:
```{r}
summary(m5 <- lmer(time.lag ~ condition*cluster + duration.tt.c + recording.no + gender + akinetic*condition + (1+condition + cluster|subj), data=df.rightedge.nospr[df.rightedge.nospr$group == "PD",], control = lmerControl(optimizer = "bobyqa")))$coef
anova(m4, m5)
```
No improvement.

Include 'condition:cluster:akinetic' interaction as fixed effect:
```{r}
summary(m6 <- lmer(time.lag ~ condition*cluster*akinetic + duration.tt.c + recording.no + gender + (1+condition + cluster|subj), data=df.rightedge.nospr[df.rightedge.nospr$group == "PD",], control = lmerControl(optimizer = "bobyqa")))$coef
anova(m4, m6)
```
No improvement.

Include 'akinetic:cluster' interaction as fixed effect:
```{r}
summary(m7 <- lmer(time.lag ~ condition*cluster + duration.tt.c + recording.no + gender + akinetic*cluster + (1+condition + cluster|subj), data=df.rightedge.nospr[df.rightedge.nospr$group == "PD",], control = lmerControl(optimizer = "bobyqa")))$coef
anova(m4, m7)
```
No improvement.