---
title: "Consonant durations in onset"
output: 
  html_document: 
    number_sections: yes
    toc: yes
    toc_float: yes
    df_print: paged
---
  
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data preprocessing

## Load required packages
```{r}
library(tidyr)
library(dplyr)
library(stringr)
library(car)
library(visreg)
library(lmerTest)
library(ggplot2)
library(mgcv)
library(itsadug)
```



## Read data
```{r save data}
df <- readRDS("C:/Users/Jidde/Desktop/analysis/data/modelling_data.rds")
df <- rename(df, onset_type = condition)
```


### Tweak data
```{r}
# Change order of levels in cluster
df$cluster <- relevel(df$cluster, ref = "sp")
df$subj <- as.factor(df$subj)
df$dur <- as.numeric(df$dur)
df <- drop_na(df, dur)
df$onset_type <- relevel(df$onset_type, ref = "C")



df.onset.sp <- df[df$seg == "p" & df$cluster == "sp",]
df.onset.sm <- df[df$seg == "m",]
df.onset.sx <- df[df$seg == "X",]

df.onset <- rbind(df.onset.sp, df.onset.sm, df.onset.sx) # drop spr cluster
rm(df.onset.sp, df.onset.sm, df.onset.sx)

table(df.onset$subj, df.onset$trial)

df.onset.s <- df[df$seg == "s" & df$onset_type == "CC",]

df.onset$onset_type <- recode_factor(df.onset$onset_type, C = "singleton", 
                                CC = "cluster")

df.onset$group <- recode_factor(df.onset$group, CTRL = "typical")
df.onset$group<- relevel(df.onset$group, ref = "PD")
```


Create ordered factor:
```{r}
df.onset$group.ord <- as.ordered(df.onset$group) 
contrasts(df.onset$group.ord) <- "contr.treatment" 

df.onset$gender.ord <- as.ordered(df.onset$gender) 
contrasts(df.onset$gender.ord) <- "contr.treatment" 
```


Create binary dummy variables:
```{r}
df.onset$IsTP <- (df.onset$group == "typical")*1
```

Visualize:
```{r}
dodge = .5
dodge <- position_dodge(width = 0.9)

ggplot(df.onset, aes(x=group, y=dur, fill=onset_type)) + geom_violin() + geom_boxplot(width=.2, position = dodge) + labs(y = "duration prevocalic consonant (ms)") + scale_fill_manual(values=c("#bfd8bf", "#bfd4ff")) + theme_bw()
dodge <- position_dodge(width = 0.9)
```


# Duration of prevocalic consonant

Fitting first model:
```{r m, cache = T}
summary(m <- bam(dur ~ s(subj, bs='re'), data=df.onset, method = "ML"))
```

Adding onset_type as fixed effect:
```{r m1, cache = T}
summary(m1 <- bam(dur ~  onset_type + s(subj, bs='re'), data=df.onset, method = "ML"))
compareML(m, m1)
```
Sign. improvement.

Add random slopes over onset_type:
```{r}
summary(m2REML <- bam(dur ~ onset_type + s(onset_type, bs = "fs", m = 1), data=df.onset , method = "REML"))
```



Adding random slopes for onset_type:
```{r m2reml, cache = T}
summary(m2REML <- bam(dur ~ onset_type + s(subj, bs='re') +  s(subj, onset_type, bs = "re"), data=df.onset , method = "REML"))
```
Sign. improvement.

Adding group as fixed effect:
```{r m3ML, cache = T}
summary(m2ML <- bam(dur ~ onset_type + s(subj, bs='re') + s(subj, onset_type, bs = "re"), data=df.onset , method = "ML"))

summary(m3ML <- bam(dur ~ group  + onset_type + s(subj, bs='re') + s(subj, onset_type, bs = "re"), data=df.onset , method = "ML"))

compareML(m2ML, m3ML)
```
No improv.

Adding interaction as fixed effect:
```{r m4ml, cache = T}
summary(m4ML <- bam(dur ~ group*onset_type + s(subj, bs='re') + s(subj, onset_type, bs = "re"), data=df.onset , method = "ML"))
compareML(m2ML, m4ML)
```
No improvement.

Add cluster
```{r m5ml, cache = T}
summary(m5ML <- bam(dur ~ onset_type + cluster + s(subj, bs='re') + s(subj, onset_type, bs = "re"), data=df.onset , method = "ML"))
compareML(m2ML, m5ML)
```
Sign. improvement.


Add random slopes for cluster:
```{r m6, cache = T}
summary(m6 <- bam(dur ~ onset_type + cluster + s(subj, bs='re') + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), data=df.onset))
```
Imrpov.

Adding onset_type*cluster as fixed effect:
```{r m7, cache = T}
summary(m6ML <- bam(dur ~ onset_type + cluster + s(subj, bs='re') + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), data=df.onset , method = "ML"))

summary(m7 <- bam(dur ~ onset_type*cluster + s(subj, bs='re') + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), data=df.onset , method = "ML"))

compareML(m6ML, m7)
```
Sign. improv.


Adding duration.tt fixed effect:
```{r m8, cache = T}
df.onset$duration.tt <- df.onset$duration.tt - mean(df.onset$duration.tt) # center continues variable
summary(m8 <- bam(dur ~ onset_type*cluster + s(duration.tt) + s(subj, bs='re') + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), method = "ML", data=df.onset))
compareML(m7, m8)
```
Sign. improvement.

Add slopes:
```{r m9, cache = T}
summary(m9 <- bam(dur ~ onset_type*cluster + s(duration.tt) + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re")  + s(duration.tt, subj, bs = "fs", m =1), data=df.onset))
```
Not required.

Check if there's a group difference:
```{r m9group, cache = T}
summary(m9group <- bam(dur ~ onset_type*cluster + s(duration.tt) + s(duration.tt, by = IsTP) + s(subj, bs = "re") + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), data=df.onset))
```
Yes.

Add appropriate random smooths:
```{r m9groslo, cache = T}
summary(m9group_slope <- bam(dur ~ onset_type*cluster + s(duration.tt) + s(duration.tt, by = IsTP) + s(subj, bs = "re") + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re") + s(duration.tt, subj, bs = "fs", m =1) + s(duration.tt, subj, by = IsTP, bs = "fs", m =1), data=df.onset))
```
Not sign. anymore.

Adding recording.no  as fixed effect:
```{r m10, cache = T}
summary(m9groupML <- bam(dur ~ group.ord + onset_type*cluster + s(duration.tt) + s(subj, bs = "re") + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), data=df.onset, method = "ML"))

summary(m10 <- bam(dur ~ group.ord + onset_type*cluster + s(duration.tt)  + s(recording.no) + s(subj, bs = "re") + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), data=df.onset, method = "ML"))

compareML(m9groupML, m10)
```
No improvement.


Recording.no separate for group?
```{r m11, cache = T}
summary(m11 <- bam(dur ~ group.ord + onset_type*cluster + s(duration.tt) + s(recording.no) + s(recording.no, by = IsTP) + s(subj, bs = "re") + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), data=df.onset))
```
No.


Adding gender as fixed effect:
```{r m12, cache = T}
summary(m11ML <- bam(dur ~ group.ord + onset_type*cluster + s(duration.tt) + s(subj, bs = "re") + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), data=df.onset, method = "ML"))

summary(m12 <- bam(dur ~ group.ord + onset_type*cluster + s(duration.tt) + gender + s(subj, bs = "re") + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), data=df.onset, method = "ML"))

compareML(m11ML, m12)
```
No sign. improvement.

In interaction with group?
```{r m12inter, cache = T}
summary(m12inter <- bam(dur ~ group.ord*gender + onset_type*cluster + s(duration.tt) + gender + s(subj, bs = "re") + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), data=df.onset, method = "ML"))
compareML(m11ML, m12inter)
```
No sign. improvement.


Adding age fixed effect:
```{r m13, cache = T}
df.onset$age <- df.onset$age - mean(df.onset$age) # center continues variable

summary(m13 <- bam(dur ~ group.ord + onset_type*cluster + s(duration.tt) + s(age) + s(subj, bs = "re") + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), data=df.onset, method = "ML"))

compareML(m11ML, m13)
```
No sign. improvement.

In interaction with group?
```{r m13inter, cache = T}
summary(m13inter <- bam(dur ~ onset_type*cluster + s(duration.tt) + s(age) + s(age, IsTP) + s(subj, bs = "re") + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), data=df.onset))
```
No sign. improvement.


Adding interaction group*cluster 
```{r m14, cache = T}
summary(m14 <- bam(dur ~ group.ord*cluster + onset_type*cluster + s(duration.tt) + s(subj, bs = "re") + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), data=df.onset, method = "ML"))

compareML(m11ML, m14)
```
No improv.

Adding interaction group*onset_type
```{r m15, cache = T}
summary(m15 <- bam(dur ~ group.ord*onset_type + onset_type*cluster + s(duration.tt) + s(subj, bs = "re") + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), data=df.onset, method = "ML"))

compareML(m11ML, m15)
```
No improv.

Add three-way interaction:
```{r m16, cache = T}
summary(m16 <- bam(dur ~ group.ord*onset_type*cluster + s(subj, bs = "re") + s(duration.tt) + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), data=df.onset, method = "ML"))

compareML(m11ML, m16)
```
Yes.


### Check assumptions

Refit using REML:
```{r m16reml, cache = T}
summary(m16REML <- bam(dur ~ group.ord*onset_type*cluster + s(subj, bs = "re") + s(duration.tt) + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), data=df.onset, method = "REML"))
```

```{r}
gam.check(m16REML)
```
Doesn't look very well, skewed right

Fit with loglink
```{r m16log, cache = T}
summary(m16REML_log <- bam(dur ~ group.ord*onset_type*cluster + s(duration.tt) + s(subj, bs = "re") + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), data=df.onset, family =Gamma(link=log)))

gam.check(m16REML_log) 
```
Still doesn't look great.

Model criticism, trimming:
```{r}
df.onset2 <- df.onset[abs(scale(resid(m16REML_log))) < 2.5, ] # trim model
(1 - (nrow(df.onset2))/nrow(df.onset)) * 100
```
3.0% removed

Refit:
```{r m16logtrim, cache = T}
summary(m16REML_log_trim <- bam(dur ~ group*onset_type*cluster + s(duration.tt) + s(subj, bs = "re") + s(subj, onset_type, bs = "re") + s(subj, cluster, bs = "re"), data=df.onset2, family =Gamma(link=log)))

gam.check(m16REML_log_trim)
```
Looks fine.

Visualize:
```{r}
plot(m16REML_log_trim, select = 1, shade = T, rug = F, xlab = "speech interval duration (ms)", ylab = "duration prevocalic C (ms)")
abline(h=0)

par(mar = c(4.5, 0.1, 0.1, 0.1)) # Set the margin on all sides to 2
plot_parametric(m16REML_log_trim, cond = list(group = "PD", cluster = c("sp")), pred=list(onset_type = c("singleton", "cluster")), main = "", xlim=c(2.5,4), xlab = "duration prevocalic consonant (ms)")
plot_parametric(m16REML_log_trim, cond = list(group = "PD", cluster = c("sm")), pred=list(onset_type = c("singleton", "cluster")), main = "", xlim=c(2.5,4), xlab = "duration prevocalic consonant (ms)")
plot_parametric(m16REML_log_trim, cond = list(group = "PD", cluster = c("sx")), pred=list(onset_type = c("singleton", "cluster")), main = "", xlim=c(2.5,4), xlab = "duration prevocalic consonant (ms)")
plot_parametric(m16REML_log_trim, cond = list(group = "typical", cluster = c("sp")), pred=list(onset_type = c("singleton", "cluster")), main = "", xlim=c(2.5,4), xlab = "duration prevocalic consonant (ms)")
plot_parametric(m16REML_log_trim, cond = list(group = "typical", cluster = c("sm")), pred=list(onset_type = c("singleton", "cluster")), main = "", xlim=c(2.5,4), xlab = "duration prevocalic consonant (ms)", ylim=c(1,-1))
plot_parametric(m16REML_log_trim, cond = list(group = "typical", cluster = c("sx")), pred=list(onset_type = c("singleton", "cluster")), main = "", xlim=c(2.5,4), xlab = "duration prevocalic consonant (ms)", ylim=c(1,-1))
```

Latex table function:
```{r}
gamtabs2 <- function (model, caption = " ", label = "tab.gam", 
    pnames = NA, snames = NA, ptab = NA, stab = NA, ...) 
{
    if (!requireNamespace("xtable", quietly = TRUE)) {
        stop("Package 'xtable' needed for this function to work. Please install it.", 
            call. = FALSE)
    }
    sum.gam <- model
    if (!inherits(model, "summary.gam")) {
        sum.gam <- summary(model)
    }
    if (is.na(ptab[1])) {
        ptab = as.data.frame(sum.gam$p.table, stringsAsFactors = FALSE)
    }
    if (is.na(stab[1])) {
        stab = as.data.frame(sum.gam$s.table, stringsAsFactors = FALSE)
    }
    if (!is.na(pnames[1])) {
        rownames(ptab) = pnames
    }
    if (!is.na(snames[1])) {
        rownames(stab) = snames
    }
    colnames(ptab)[4] = "p-value"
    colnames(ptab)[3] = "t-value"
    ptab.cnames = colnames(ptab)
    stab.cnames = colnames(stab)
    stab.cnames[3] = "F-value"
    colnames(ptab) = c("A", "B", "C", "D")
    if (ncol(stab) != 0) {
        colnames(stab) = colnames(ptab)
    }
    tab = rbind(ptab, stab)
    colnames(tab) = ptab.cnames
    tab = round(tab, 4)
    m = data.frame(matrix(0, nrow(tab), ncol(tab)), stringsAsFactors = FALSE)
    for (i in 1:nrow(tab)) {
        for (j in 1:4) {
            if ((j == 4) & (tab[i, j] < 1e-03)) {
                m[i, j] = "< 0.001"
            }
            else {
                m[i, j] = sprintf("%3.3f", tab[i, j])
            }
        }
    }
    colnames(m) = colnames(tab)
    rownames(m) = rownames(tab)
    tab = m
    tab2 = rbind(c(ptab.cnames), tab[1:nrow(ptab), ])
    if (nrow(stab) > 0) {
        tab2 = rbind(tab2, c(stab.cnames), tab[(nrow(ptab) + 
            1):nrow(tab), ])
    }
    if (nrow(stab)) {
        rownames(tab2)[(nrow(ptab) + 2)] = "B. smooth terms"
    }
    rownames(tab2)[1] = "A. parametric coefficients"
    for (i in 1:nrow(tab2)) {
        if (tab2[i, 4] == "0") 
            tab2[i, 4] = "< 0.0001"
        if (length(grep("\\.", tab2[i, 2])) == 0) 
            tab2[i, 2] = paste(tab2[i, 2], ".0000", sep = "")
    }
    print(xtable::xtable(tab2, caption = caption, label = label, 
        align = "lrrrr"), include.colnames = FALSE, hline.after = c(0, 
        (nrow(ptab) + 1), nrow(tab2)), ...)
}
```

Latex table:
```{r}
options(xtable.comment = FALSE)
gamtabs2(m16REML_log_trim)
```

# Session info
```{r}
sessionInfo()
```

