---
title: "Relationship between IPI and duration of C1/C2"
output: 
  html_document: 
    number_sections: yes
    toc: yes
    toc_float: yes
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data preprocessing

## Load required packages
```{r}
library("tidyr")
library("dplyr")
library("stringr")
library("car")
library("ggplot2")
library("mgcv")
library("itsadug")
```


## Read data
```{r save data}
df <- readRDS("C:/Users/Jidde/Desktop/analysis/data/modelling_data_CC.rds")
df <- rename(df, IPI = time.lag.norm)
```

## Clean data
Remove all the rows that have NA in the dependent variable and drop 'empty' levels, convert variables to factors and create trimmed dataset.
```{r remove NA}
df$subj <- as.factor(df$subj)

df <- df[complete.cases(df$IPI),]
df <- droplevels(df)

# Sort data
df <- df[order(df$subj, df$recording.no),]

# convert to factors/numeric
df$prompt <- as.factor(df$prompt)
df$condition <- as.factor(df$condition)


# remove all the recordings without a speech rate measure:
# df <- df %>% drop_na(duration.tt)

# remove participants with only a few datapoints
#remove <- c("PD19", "CTRL19")
#df.trim <- df[! df$subj %in% remove, ]
#df.trim <- droplevels(df.trim)


# create df without sx
#df.trim <- df.trim[! df.trim$cluster == "sx", ]
#df.trim <- droplevels(df.trim)
#levels(df.trim$cluster)

# create df without spr
df.nospr <- df[! df$cluster == "spr", ]
df.nospr <- droplevels(df.nospr)
levels(df.nospr$cluster)
df.nospr$cluster <- relevel(df.nospr$cluster, ref = "sm")



# merge /sm/ and /sp/
df.nospr$c2type <- ifelse(df.nospr$cluster == "sx","lingual C2","labial C2")
df.nospr$c2type <- as.factor(df.nospr$c2type)
table(df.nospr$c2type, df.nospr$cluster)

df.nospr$condition <- recode_factor(df.nospr$condition, C = "singleton", 
                                CC = "cluster")

df.nospr$group <- recode_factor(df.nospr$group, CTRL = "typical")
df.nospr$group<- relevel(df.nospr$group, ref = "PD")


rm(df)
```

Create one var for duration of C1/C2
```{r}
df.nospr <- df.nospr %>%
  gather(whichC, durationC, dur.C1.norm, dur.C2.norm) %>%
  arrange(subj, group, recording.no)

df.nospr$whichC <- as.factor(df.nospr$whichC)

#df.nospr$durationC <- df.nospr$durationC - mean(df.nospr$durationC) # center continues variable
df.nospr$cluster <- relevel(df.nospr$cluster, ref = "sx")
df.nospr$whichC <- relevel(df.nospr$whichC, ref = "dur.C1.norm")
df.nospr$whichC <- recode_factor(df.nospr$whichC, dur.C1.norm = "initial", 
                                dur.C2.norm = "prevocalic")
```

Create binary dummy variables:
```{r}
df.nospr$IsC2 <- (df.nospr$whichC == "prevocalic")*1
df.nospr$IsTP <- (df.nospr$group == "typical")*1
df.nospr$IsSm = (df.nospr$cluster == "sm")*1
df.nospr$IsSx = (df.nospr$cluster == "sx")*1
df.nospr$IsC2lin = (df.nospr$cluster == "sx")*1

df.nospr$IsTP_C2 = (df.nospr$whichC == "prevocalic" & df.nospr$group == "typical")*1
df.nospr$IsPD_C2 = (df.nospr$whichC == "prevocalic" & df.nospr$group == "PD")*1

df.nospr$IsTP_C2lin = (df.nospr$cluster == "sx" & df.nospr$group == "typical")*1
```


Create ordered factor:
```{r}
df.nospr$group.ord <- as.ordered(df.nospr$group) 
contrasts(df.nospr$group.ord) <- "contr.treatment" 

df.nospr$whichC.ord <- as.ordered(df.nospr$whichC) 
contrasts(df.nospr$whichC.ord) <- "contr.treatment" 

df.nospr$cluster.ord <- as.ordered(df.nospr$cluster) 
contrasts(df.nospr$cluster.ord) <- "contr.treatment" 

df.nospr$gender.ord <- as.ordered(df.nospr$gender) 
contrasts(df.nospr$gender.ord) <- "contr.treatment" 

df.nospr$c2type.ord <- as.ordered(df.nospr$c2type) 
contrasts(df.nospr$c2type.ord) <- "contr.treatment" 

df.nospr$TPlingual.ord <- as.ordered(df.nospr$group =="typical" & df.nospr$c2type == "lingual C2") 
contrasts(df.nospr$TPlingual.ord) <- "contr.treatment" 

df.nospr$PDlingual.ord <- as.ordered(df.nospr$group =="PD" & df.nospr$c2type == "lingual C2") 
contrasts(df.nospr$TPlingual.ord) <- "contr.treatment" 

df.nospr$c2type_group <- interaction(df.nospr$c2type, df.nospr$group)
df.nospr$c2type_group.ord <- as.ordered(df.nospr$c2type_group) 
contrasts(df.nospr$c2type_group.ord) <- "contr.treatment" 
```


## Hypothesis testing

Base model:
```{r}
summary(mBase <- bam(durationC ~ s(subj, bs = "re"), data = df.nospr, method = "ML"))
```
OK.

Add smooth over IPI
```{r}
summary(m1 <- bam(durationC ~ s(IPI) + s(subj, bs = "re"), data = df.nospr, method = "ML"))
compareML(mBase, m1)
```
Yes.

Add random smooth?
```{r}
summary(m2 <- bam(durationC ~ s(IPI) + s(IPI, subj, bs = "fs", m =1), data = df.nospr, discrete = T, nthreads = 2))
```
Yes.

Plot
```{r hypo, cache = T}
plot_smooth(m2, view = "IPI")
```

Check:
```{r}
gam.check(m2)
```

Looks heavily right skewed.

Fit with log:
```{r hypolog, cache = T}
summary(m2_log <- bam(durationC ~ s(IPI) + s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), discrete = T, nthreads = 2))

gam.check(m2_log)
```
Looks a bit better.

Compare model with and without IPI using the right (log) distribution:
```{r compareIPI, cache = T}
summary(mBaseLog <- bam(durationC ~ s(subj, bs = "re"), data = df.nospr, family =Gamma(link=log), method = "ML"))

summary(mHypolog <- bam(durationC ~ s(IPI) + s(subj, bs = "re"), data = df.nospr, family =Gamma(link=log), method = "ML"))

compareML(mBaseLog, mHypolog)
```



Model criticism, trimming:
```{r }
df.nospr2 <- df.nospr[abs(scale(resid(m2_log))) < 2.5, ] # trim model
(1 - (nrow(df.nospr2))/nrow(df.nospr)) * 100
```
0.5% removed.

Refit:
```{r hypotrim, cache = T}
summary(hypo_log_trim <- bam(durationC ~ s(IPI) + s(IPI, subj, bs = "fs", m =1), data = df.nospr2, family =Gamma(link=log), discrete = T, nthreads = 2))
gam.check(hypo_log_trim)
```
OK.

Visualize:
```{r}
#Plot using ordered factor:
summary(mplot <- bam(durationC ~ group.ord + s(IPI) + s(IPI, subj, bs = "fs", m =1), data = df.nospr2, family =Gamma(link=log), discrete = T, nthreads = 2))

plot(mplot, select = 1, shade = T, rug = F, ylab = "duration consonant (ms)", xlab = "IPI (ms)")
abline(h=0)
```
# GAM models

## Exploraratory testing

Add group:
```{r ex0, cache = T}
summary(m2_log.alt <- bam(durationC ~ s(IPI) + s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), select = T, discrete = T, nthreads = 2))

summary(ex0.alt <- bam(durationC ~ group + s(IPI) + s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), select = T, discrete = T, nthreads = 2))

compareML(m2_log.alt, ex0.alt)
```


Binary smooth for each group over IPI?
```{r exgroupsm, cache = T}
summary(mGroupBin<- bam(durationC ~ s(IPI) + s(IPI, by = IsTP) + s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), discrete = T, nthreads = 2))
```



Add whichC (compare using select = T)
```{r ex1, cache = T}
summary(ex1.alt <- bam(durationC ~ s(IPI) + whichC + s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), select = T, discrete = T, nthreads = 2))

compareML(m2_log.alt, ex1.alt)
```
Yes.


Add slopes
```{r ex2, cache = T}
summary(ex2 <- bam(durationC ~ s(IPI) + whichC + s(subj, whichC, bs="re") + s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), discrete = T, nthreads = 2))
```
Required.


Add cluster:
```{r ex4, cache = T}
summary(ex2.alt <- bam(durationC ~ s(IPI) + whichC + s(subj, whichC, bs="re") + s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), select = T, discrete = T, nthreads = 2))

summary(ex3.alt <- bam(durationC ~ s(IPI) + whichC + cluster + s(subj, whichC, bs="re") + s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), select = T, discrete = T, nthreads = 2))

compareML(ex2.alt, ex3.alt)
```
No.


In interaction with group?
```{r}
summary(ex4.alt <- bam(durationC ~ s(IPI) + whichC + group*cluster + s(subj, whichC, bs="re") + s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), select = T, discrete = T, nthreads = 2))

compareML(ex2.alt, ex4.alt)
```
No.


Add c2type?
```{r ex5alt, cache = T}
summary(ex5.alt <- bam(durationC ~ s(IPI) + whichC + c2type + s(subj, whichC, bs="re") + s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), select = T, discrete = T, nthreads = 2))

compareML(ex2.alt, ex5.alt)
```
No.


In interaction with group?
```{r ex6alt, cache = T}
summary(ex6.alt <- bam(durationC ~ s(IPI) + whichC + group*c2type + s(subj, whichC, bs="re") + s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), select = T, discrete = T, nthreads = 2))

compareML(ex2.alt, ex6.alt)
```
No.


Add recording.no:
```{r ex9, cache = T}
summary(ex5.alt <- bam(durationC ~ s(IPI) + whichC + s(subj, whichC, bs="re") + s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), select = T, discrete = T, nthreads = 2))

summary(ex6.alt <- bam(durationC ~ s(IPI) + whichC + s(recording.no) + s(subj, whichC, bs="re") + s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), select = T, discrete = T, nthreads = 2))

compareML(ex5.alt, ex6.alt)
```
No.


Separate for group?
```{r ex11, cache = T}
summary(ex7 <- bam(durationC ~ s(IPI) + whichC + s(recording.no) + s(recording.no, by = IsTP) + s(subj, whichC, bs="re") + s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), discrete = T, nthreads = 2))
```
No.


Add gender:
```{r ex12, cache = T}
summary(ex8.alt <- bam(durationC ~ s(IPI) + whichC + gender + s(subj, whichC, bs="re") +  s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), select = T, discrete = T, nthreads = 2))

compareML(ex5.alt, ex8.alt)
```
No.


Add gender in interaction with group?
```{r ex13, cache = T}
summary(ex9.alt <- bam(durationC ~ s(IPI) + whichC + group*gender + s(subj, whichC, bs="re") + s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), select = T, discrete = T, nthreads = 2))

compareML(ex5.alt, ex9.alt)
```
No.


Add age:
```{r ex14, cache = T}
summary(ex10.alt <- bam(durationC ~ s(IPI) + whichC  + s(age) + s(subj, whichC, bs="re") + s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), select = T, discrete = T, nthreads = 2))

compareML(ex5.alt, ex10.alt)
```
No improvement.


In interaction with group:
```{r ex15, cache = T}
summary(ex10.alt <- bam(durationC ~ s(IPI) + whichC + s(age) + s(age, by = IsTP) + s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), discrete = T, nthreads = 2))
```
No improvement.


Check once more if we need IPI for group?
```{r ex16, cache=T }
summary(ex11 <- bam(durationC ~ s(IPI) + s(IPI, by = IsTP) + whichC + s(subj, whichC, bs="re") + s(IPI, subj, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), discrete = T, nthreads = 2))
```
No.


Separate smooth over IPI for each level of whichC (including appropriate random smooth)?
```{r exsep, cache=T}
summary(ex11 <- bam(durationC ~ s(IPI) + s(IPI, by = IsC2) + s(IPI, subj, bs = "fs", m =1) + s(IPI, subj, by = whichC.ord, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), discrete = T, nthreads = 2))
```
Yes.


Separate per group?
```{r exsepgroup, cache=T}
summary(ex12 <- bam(durationC ~ group + s(IPI, by = group) + s(IPI, by = IsTP_C2) + s(IPI, by = IsPD_C2) + s(IPI, subj, bs = "fs", m =1) + s(IPI, subj, by = whichC.ord, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), discrete = T, nthreads = 2))
```
Both groups distinguish C1 from C2.


Check if this there as a difference between groups:
```{r exdiffsepgroup, cache=T}
summary(ex12 <- bam(durationC ~ group + s(IPI, by = group) + s(IPI, by = IsC2) + s(IPI, by = IsTP_C2)  + s(IPI, subj, bs = "fs", m =1) + s(IPI, subj, by = whichC.ord, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), discrete = T, nthreads = 2))
```
No difference between the groups.


Check once more whether we need cluster:
```{r ex13cluster, cache=T}
summary(ex11.alt <- bam(durationC ~ s(IPI) + s(IPI, by = IsC2) + s(IPI, subj, bs = "fs", m =1) + s(IPI, subj, by = whichC.ord, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), select = T, discrete = T, nthreads = 2))

summary(ex13.alt <- bam(durationC ~ s(IPI) + s(IPI, by = IsC2) + cluster + s(IPI, subj, bs = "fs", m =1) + s(IPI, subj, by = whichC.ord, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), select = T, discrete = T, nthreads = 2))

compareML(ex11.alt, ex13.alt)
```
No.


In interaction with group?
```{r ex14cluster, cache=T}
summary(ex14.alt <- bam(durationC ~ s(IPI) + s(IPI, by = IsC2) + cluster*group + s(IPI, subj, bs = "fs", m =1) + s(IPI, subj, by = whichC.ord, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), select = T, discrete = T, nthreads = 2))

compareML(ex11.alt, ex14.alt)
```
No.


Check once more whether we need c2type:
```{r ex15.alt, cache=T}
summary(ex15.alt <- bam(durationC ~ s(IPI) + s(IPI, by = IsC2) + c2type + s(IPI, subj, bs = "fs", m =1) + s(IPI, subj, by = whichC.ord, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), select = T, discrete = T, nthreads = 2))

compareML(ex11.alt, ex15.alt)
```
No.


Check once more whether we need c2type x group:
```{r ex16.alt, cache=T}
summary(ex16.alt <- bam(durationC ~ group*c2type + s(IPI) + s(IPI, by = IsC2) + s(IPI, subj, bs = "fs", m =1) + s(IPI, subj, by = whichC.ord, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), select = T, discrete = T, nthreads = 2))

compareML(ex11.alt, ex16.alt)
```
Yes.


Add slopes?
```{r ex17slop, cache = T}
summary(ex17 <- bam(durationC ~ group*c2type + s(IPI) + s(IPI, by = IsC2) + s(subj, c2type, bs = "re") + s(IPI, subj, bs = "fs", m =1) + s(IPI, subj, by = whichC.ord, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), discrete = T, nthreads = 2))
```
Yes.


Final model using ordered factors (in order to distinguish constant from non-linear diff:
```{r finalord, cache=T} 
summary(Final.ord <- bam(durationC ~ group*c2type + whichC.ord + s(IPI) + s(IPI, by = whichC.ord)  + s(subj, c2type, bs = "re") + s(IPI, subj, bs = "fs", m =1) + s(IPI, subj, by = whichC.ord, bs = "fs", m =1), data = df.nospr, family =Gamma(link=log), discrete = T, nthreads = 2))
```


# Check assumptions

Fit ordered factor model without separate smooths over IPI per group:
```{r exFinal, cache = T}
mFinal.ord <- Final.ord
```


Check:
```{r}
gam.check(mFinal.ord)
```
Fine.

Model criticism, trimming:
```{r}
df.nospr3 <- df.nospr[abs(scale(resid(mFinal.ord))) < 2.5, ] # trim model
(1 - (nrow(df.nospr3))/nrow(df.nospr)) * 100
```
2.7% gets removed.

Refit:
```{r exFinaltrim, cache = T}
summary(Final.ord_trim <- bam(durationC ~ group*c2type + whichC.ord + s(IPI) + s(IPI, by = whichC.ord)  + s(subj, c2type, bs = "re") + s(IPI, subj, bs = "fs", m =1) + s(IPI, subj, by = whichC.ord, bs = "fs", m =1), data = df.nospr3, family =Gamma(link=log), discrete = T, nthreads = 2))
```
same pattern

Visualize:
```{r}
plot_parametric(Final.ord_trim, cond = list(group = "PD"), pred=list(c2type = c("labial C2", "lingual C2")), main = "", xlab = "duration consonants (ms)", xlim=c(-4.1,-3.4))

plot_parametric(Final.ord_trim, cond = list(group = "typical"), pred=list(c2type = c("labial C2", "lingual C2")), main = "", xlab = "duration consonants (ms)", xlim=c(-4.1,-3.4))


plot_smooth(Final.ord_trim, view="IPI", plot_all="whichC.ord", rm.ranef=TRUE, xlab = "IPI (ms)", col=c("red", "darkgreen"), ylab = "duration consonant (ms)", rug=F)
```


Latex table function:
```{r}
gamtabs2 <- function (model, caption = " ", label = "tab.gam", 
    pnames = NA, snames = NA, ptab = NA, stab = NA, ...) 
{
    if (!requireNamespace("xtable", quietly = TRUE)) {
        stop("Package 'xtable' needed for this function to work. Please install it.", 
            call. = FALSE)
    }
    sum.gam <- model
    if (!inherits(model, "summary.gam")) {
        sum.gam <- summary(model)
    }
    if (is.na(ptab[1])) {
        ptab = as.data.frame(sum.gam$p.table, stringsAsFactors = FALSE)
    }
    if (is.na(stab[1])) {
        stab = as.data.frame(sum.gam$s.table, stringsAsFactors = FALSE)
    }
    if (!is.na(pnames[1])) {
        rownames(ptab) = pnames
    }
    if (!is.na(snames[1])) {
        rownames(stab) = snames
    }
    colnames(ptab)[4] = "p-value"
    colnames(ptab)[3] = "t-value"
    ptab.cnames = colnames(ptab)
    stab.cnames = colnames(stab)
    stab.cnames[3] = "F-value"
    colnames(ptab) = c("A", "B", "C", "D")
    if (ncol(stab) != 0) {
        colnames(stab) = colnames(ptab)
    }
    tab = rbind(ptab, stab)
    colnames(tab) = ptab.cnames
    tab = round(tab, 4)
    m = data.frame(matrix(0, nrow(tab), ncol(tab)), stringsAsFactors = FALSE)
    for (i in 1:nrow(tab)) {
        for (j in 1:4) {
            if ((j == 4) & (tab[i, j] < 1e-03)) {
                m[i, j] = "< 0.001"
            }
            else {
                m[i, j] = sprintf("%3.3f", tab[i, j])
            }
        }
    }
    colnames(m) = colnames(tab)
    rownames(m) = rownames(tab)
    tab = m
    tab2 = rbind(c(ptab.cnames), tab[1:nrow(ptab), ])
    if (nrow(stab) > 0) {
        tab2 = rbind(tab2, c(stab.cnames), tab[(nrow(ptab) + 
            1):nrow(tab), ])
    }
    if (nrow(stab)) {
        rownames(tab2)[(nrow(ptab) + 2)] = "B. smooth terms"
    }
    rownames(tab2)[1] = "A. parametric coefficients"
    for (i in 1:nrow(tab2)) {
        if (tab2[i, 4] == "0") 
            tab2[i, 4] = "< 0.0001"
        if (length(grep("\\.", tab2[i, 2])) == 0) 
            tab2[i, 2] = paste(tab2[i, 2], ".0000", sep = "")
    }
    print(xtable::xtable(tab2, caption = caption, label = label, 
        align = "lrrrr"), include.colnames = FALSE, hline.after = c(0, 
        (nrow(ptab) + 1), nrow(tab2)), ...)
}
```

Latex table:
```{r}
options(xtable.comment = FALSE)
gamtabs2(Final.ord_trim)
```


# Session info
```{r}
sessionInfo()
```

