---
title: "C-center RSD models"
output: 
  html_document: 
    number_sections: yes
    toc: yes
    toc_float: yes
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preprocessing

## Load required packages
```{r}
library(tidyr)
library(dplyr)
library(stringr)
library(lme4)
library(car)
library(visreg)
library(ggplot2)
library(mgcv)
library(itsadug)
```



## Read data
```{r save data}
df <- readRDS("C:/Users/Jidde/Desktop/analysis/data/modelling_data_RSD.rds")
df <- rename(df, interval_type = condition)
```


## Tweak data
```{r}
# Change order of levels in cluster
df$cluster <- factor(df$cluster, levels = c("sp", "sm", "sx", "spr"))
df <- subset(df, RSD != 0)

# create df without spr
df<- df[! df$cluster == "spr", ]
df <- droplevels(df)
levels(df$cluster)
df$group <- recode_factor(df$group, CTRL = "typical")
df$group<- relevel(df$group, ref = "PD")

```

Create ordered factors:
```{r}
df$group.ord <- as.ordered(df$group) 
contrasts(df$group.ord) <- "contr.treatment" 
```


Create binary dummy variables:
```{r}
df$IsTP <- (df$group == "typical")*1
```


Visualize:
```{r}
dodge = .5

ggplot(df, aes(x=interval_type, y=RSD, fill=interval_type)) + geom_violin()  + geom_boxplot(width=.2) + labs(y = "relative standard deviation (RSD)")  + scale_fill_manual(values=c("#bfd8bf", "#ffbfbf", "#bfd4ff")) + theme_bw()

ggplot(df[df$group =="typical",], aes(x=interval_type, y=RSD, fill=interval_type)) + geom_violin()  + geom_boxplot(width=.2) + labs(y = "relative standard deviation (RSD)")  + scale_fill_manual(values=c("#bfd8bf", "#ffbfbf", "#bfd4ff")) + theme_bw()

ggplot(df[df$group =="PD",], aes(x=interval_type, y=RSD, fill=interval_type)) + geom_violin()  + geom_boxplot(width=.2) + labs(y = "relative standard deviation (RSD)")  + scale_fill_manual(values=c("#bfd8bf", "#ffbfbf", "#bfd4ff")) + theme_bw()
```

# GAM models for RSD analysis

## Hypothesis test
Base model:
```{r hypo, cache =T}
summary(mhypo <- bam(RSD ~ s(subj, bs='re'), data=df, method = "ML"))
```

Add interval_type?
```{r mhypointervaltype, cache =T}
summary(mhypointerval_type <- bam(RSD ~ interval_type + s(subj, bs='re'), data=df, method = "ML"))
compareML(mhypo, mhypointerval_type)
```
Yes.


Random slopes?
```{r mhyposlopes, cache =T}
summary(mhypoSlopes <- bam(RSD ~ interval_type + s(subj, bs='re') + s(subj, interval_type, bs='re'), data=df, method = "REML"))
```
No.


Scaled t?
```{r mhypo2_scat, cache =T}
summary(mhypo2_scat <- bam(RSD ~ interval_type + s(subj, bs='re'), data=df, method ="fREML", family = "scat", discrete = T))
gam.check(mhypo2_scat)
```

Log?
```{r mhypo2log, cache =T}
summary(mhypo2_log <- bam(RSD ~ interval_type + s(subj, bs='re'), data=df, family =Gamma(link=log)))
gam.check(mhypo2_log)
```
Looks best.

Model criticism, trimming:
```{r}
df2 <- df[abs(scale(resid(mhypo2_log))) < 2.5, ] # trim model
(1 - (nrow(df2))/nrow(df)) * 100
```
2.2% removed.


Refit:
```{r mhypo2logtrim, cache =T}
summary(mhypo2_log_trim <- bam(RSD ~ interval_type +s(subj, bs='re'), data=df2, family =Gamma(link=log)))
gam.check(mhypo2_log_trim)
```

Plot:
```{r}
plot_parametric(mhypo2_log_trim, pred=list(interval_type = c("left edge", "C-center", "right edge")), main = "", xlab = "relative standard deviation (RSD)", xlim=c(1.5,3.5))
```


## Exploratory analysis

Add group?
```{r mgroup, cache = T}
summary(mhypo2_log_trim_ML <- bam(RSD ~ interval_type + s(subj, bs='re'), data=df, family =Gamma(link=log), method = "ML"))

summary(mgroup <- bam(RSD ~ group + interval_type + s(subj, bs='re'), data=df, family =Gamma(link=log), method = "ML"))
compareML(mhypo2_log_trim_ML, mgroup)
```
No.


Add group in interaction with interval type?
```{r mgroupinter, cache = T}
summary(mgroupinter <- bam(RSD ~ group*interval_type + s(subj, bs='re'), data=df, family =Gamma(link=log), method = "ML"))
compareML(mhypo2_log_trim_ML, mgroupinter)
```
No.


Adding cluster as fixed effect:
```{r ex1, cache =T}
summary(ex1 <- bam(RSD ~ s(subj, bs='re') + cluster + interval_type, data=df, family =Gamma(link=log), method = "ML"))
compareML(mhypo2_log_trim_ML, ex1)
```
Sign. improvement.


Adding slopes:
```{r ex2, cache =T}
summary(ex2 <- bam(RSD ~   interval_type + cluster + s(subj, bs='re') +  s(subj, cluster, bs='re'), data=df, family =Gamma(link=log)))
```
Yes.


Check once more if we need slopes for interval_type
```{r excondslopes, cache =T}
summary(ex_cond_slopes <- bam(RSD ~ interval_type + cluster + s(subj, bs='re') +  s(subj, interval_type, bs='re') +  s(subj, cluster, bs='re'), data=df, family =Gamma(link=log)))
```
We do.


Adding interval_type*cluster interaction:
```{r ex4, cache =T}
summary(ex3 <- bam(RSD ~   interval_type + cluster + s(subj, bs='re') + s(subj, interval_type, bs='re') + s(subj, cluster, bs='re'), data=df, family =Gamma(link=log), method = "ML"))

summary(ex4 <- bam(RSD ~   interval_type + interval_type*cluster + s(subj, bs='re') +  s(subj, interval_type, bs='re') +  s(subj, cluster, bs='re'), data=df, family =Gamma(link=log), method = "ML"))
compareML(ex3,ex4)
```
Sign. improvement.


Add speech rate
```{r ex5, cache =T}
summary(ex5 <- bam(RSD ~ interval_type + interval_type*cluster + s(mean.sr) + s(subj, bs='re') +  s(subj, interval_type, bs='re')+  s(subj, cluster, bs='re'), data=df, family =Gamma(link=log), method = "ML"))
compareML(ex4, ex5)
```
No improv.


Add speech rate separate for group?
```{r ex6, cache =T}
summary(ex6 <- bam(RSD ~   group.ord + interval_type + interval_type*cluster + s(mean.sr, by = IsTP) + s(subj, bs='re') +  s(subj, interval_type, bs='re') + s(subj, cluster, bs='re'), data=df, family =Gamma(link=log)))
```
No.


Add gender
```{r ex7, cache =T}
summary(ex7 <- bam(RSD ~   interval_type + interval_type*cluster + gender + s(subj, bs='re')+  s(subj, interval_type, bs='re') +  s(subj, cluster, bs='re'), data=df, family =Gamma(link=log), method = "ML"))
compareML(ex4, ex7)
```
No sign. improvement.


In interaction with group?
```{r ex7inter, cache =T}
summary(ex7inter <- bam(RSD ~   interval_type + interval_type*cluster + gender*group + s(subj, bs='re')+  s(subj, interval_type, bs='re') +  s(subj, cluster, bs='re'), data=df, family =Gamma(link=log), method = "ML"))
compareML(ex4, ex7inter)
```
No improv.


Add age
```{r ex8, cache =T}
summary(ex8 <- bam(RSD ~   interval_type + interval_type*cluster + s(age) + s(subj, bs='re')+  s(subj, interval_type, bs='re') +  s(subj, cluster, bs='re'), data=df, family =Gamma(link=log), method = "ML"))
compareML(ex4, ex8)
```
Becomes worse.


In interaction with group?
```{r ex8inter, cache =T}
summary(ex8inter <- bam(RSD ~   interval_type + interval_type*cluster + s(age, by = IsTP) + s(subj, bs='re')+  s(subj, interval_type, bs='re') +  s(subj, cluster, bs='re'), data=df, family =Gamma(link=log)))
```
No improv.


Add interaction cluster x group
```{r ex10, cache =T}
summary(ex10 <- bam(RSD ~  interval_type + group*cluster + interval_type*cluster + s(subj, bs='re') +  s(subj, interval_type, bs='re')+  s(subj, cluster, bs='re'), data=df, family =Gamma(link=log), method = "ML"))
compareML(ex4, ex10)
```
No sign. improvement.


Add threeway interaction:
```{r ex11, cache =T}
summary(ex11 <- bam(RSD ~  group*interval_type*cluster + s(subj, bs='re') +  s(subj, interval_type, bs='re')+  s(subj, cluster, bs='re'), data=df, family =Gamma(link=log), method = "ML"))
compareML(ex4, ex11)
```
No.


Check if we need interaction between interval_type and group?
```{r ex12, cache =T}
summary(ex12 <- bam(RSD ~   interval_type*group + interval_type*cluster + s(subj, bs='re') +  s(subj, interval_type, bs='re') +  s(subj, cluster, bs='re'), data=df, family =Gamma(link=log), method = "ML"))

compareML(ex4, ex12)
```
No, so we leave it out in the final model.


Add cluster in interaction with group:
```{r ex13, cache =T}
summary(ex13 <- bam(RSD ~ interval_type + interval_type*cluster + group*cluster + s(subj, bs='re') +  s(subj, interval_type, bs='re') +  s(subj, cluster, bs='re'), data=df, family =Gamma(link=log), method = "ML"))

compareML(ex4, ex13)
```



## Checking assumptions

Fit final model using REML and check assumptions:
```{r ex5reml, cache =T}
summary(ex5REML <- bam(RSD ~   interval_type*cluster + s(subj, bs='re')+  s(subj, interval_type, bs='re') +  s(subj, cluster, bs='re'), data=df, family =Gamma(link=log)))
gam.check(ex5REML)
```


Model criticism, trimming:
```{r}
df3 <- df[abs(scale(resid(ex5REML))) < 2.5, ] # trim model
(1 - (nrow(df3))/nrow(df)) * 100
```
2.2% data removed.


Refit:
```{r ex5remltrim, cache =T}
summary(ex5REML_trim <- bam(RSD ~   interval_type*cluster + s(subj, bs='re') +  s(subj, cluster, bs='re')+  s(subj, interval_type, bs='re') , data=df3, family =Gamma(link=log)))
gam.check(ex5REML_trim)
```
Same pattern.


Visualize:
```{r}
plot_parametric(ex5REML_trim, cond = list(cluster = "sp"), pred=list(interval_type = c("left edge", "C-center", "right edge")), main = "", xlab = "relative standard deviation (RSD)", xlim=c(1.5,3.5))
plot_parametric(ex5REML_trim, cond = list(cluster = "sm"), pred=list(interval_type = c("left edge", "C-center", "right edge")), main = "", xlab = "relative standard deviation (RSD)", xlim=c(1.5,3.5))
plot_parametric(ex5REML_trim, cond = list(cluster = "sx"), pred=list(interval_type = c("left edge", "C-center", "right edge")), main = "", xlab = "relative standard deviation (RSD)", xlim=c(1.5,3.5))
```

# Latex table

Latex table function:
```{r}
gamtabs2 <- function (model, caption = " ", label = "tab.gam", 
    pnames = NA, snames = NA, ptab = NA, stab = NA, ...) 
{
    if (!requireNamespace("xtable", quietly = TRUE)) {
        stop("Package 'xtable' needed for this function to work. Please install it.", 
            call. = FALSE)
    }
    sum.gam <- model
    if (!inherits(model, "summary.gam")) {
        sum.gam <- summary(model)
    }
    if (is.na(ptab[1])) {
        ptab = as.data.frame(sum.gam$p.table, stringsAsFactors = FALSE)
    }
    if (is.na(stab[1])) {
        stab = as.data.frame(sum.gam$s.table, stringsAsFactors = FALSE)
    }
    if (!is.na(pnames[1])) {
        rownames(ptab) = pnames
    }
    if (!is.na(snames[1])) {
        rownames(stab) = snames
    }
    colnames(ptab)[4] = "p-value"
    colnames(ptab)[3] = "t-value"
    ptab.cnames = colnames(ptab)
    stab.cnames = colnames(stab)
    stab.cnames[3] = "F-value"
    colnames(ptab) = c("A", "B", "C", "D")
    if (ncol(stab) != 0) {
        colnames(stab) = colnames(ptab)
    }
    tab = rbind(ptab, stab)
    colnames(tab) = ptab.cnames
    tab = round(tab, 4)
    m = data.frame(matrix(0, nrow(tab), ncol(tab)), stringsAsFactors = FALSE)
    for (i in 1:nrow(tab)) {
        for (j in 1:4) {
            if ((j == 4) & (tab[i, j] < 1e-03)) {
                m[i, j] = "< 0.001"
            }
            else {
                m[i, j] = sprintf("%3.3f", tab[i, j])
            }
        }
    }
    colnames(m) = colnames(tab)
    rownames(m) = rownames(tab)
    tab = m
    tab2 = rbind(c(ptab.cnames), tab[1:nrow(ptab), ])
    if (nrow(stab) > 0) {
        tab2 = rbind(tab2, c(stab.cnames), tab[(nrow(ptab) + 
            1):nrow(tab), ])
    }
    if (nrow(stab)) {
        rownames(tab2)[(nrow(ptab) + 2)] = "B. smooth terms"
    }
    rownames(tab2)[1] = "A. parametric coefficients"
    for (i in 1:nrow(tab2)) {
        if (tab2[i, 4] == "0") 
            tab2[i, 4] = "< 0.0001"
        if (length(grep("\\.", tab2[i, 2])) == 0) 
            tab2[i, 2] = paste(tab2[i, 2], ".0000", sep = "")
    }
    print(xtable::xtable(tab2, caption = caption, label = label, 
        align = "lrrrr"), include.colnames = FALSE, hline.after = c(0, 
        (nrow(ptab) + 1), nrow(tab2)), ...)
}
```

Latex table:
```{r}
options(xtable.comment = FALSE)
gamtabs2(ex5REML_trim)
```


# Session info
```{r}
sessionInfo()
```
